---
title: "01-temp-conditions"
author: "Yaamini Venkataraman"
date: "8/18/2022"
output: html_document
---

# Set up R Markdown document

```{bash}
mkdir ../output/01-temp-conditions #Make output directory
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "../output/01-temp-conditions/") #Set root directory
```

#Install packages

```{r}
#install.packages("tidyverse")
require(tidyverse)
```

```{r}
sessionInfo()
```

# Import and format data

Outliers from when HOBO loggers recorded data while outside of the tank were removed manually.

```{r}
#Import CSV output from HOBO loggers. Skip first line of CSV and specify header. Retain second and third columns, rename columns, and format dateTime column correctly

tank1a <- read.csv("../data/HOBO/yrv_1.csv", skip = 1, header = TRUE) %>% select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp = Temp...C..LGR.S.N..21435404..SEN.S.N..21435404.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%r"))
tank1b <- read.csv("../data/HOBO/yrv_2.csv", skip = 1, header = TRUE) %>% select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp = Temp...C..LGR.S.N..21435401..SEN.S.N..21435401.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%r"))

tank2a <- read.csv("../data/HOBO/yrv_3.csv", skip = 1, header = TRUE) %>% select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp = Temp...C..LGR.S.N..21435198..SEN.S.N..21435198.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%r"))
tank2b <- read.csv("../data/HOBO/yrv_4.csv", skip = 1, header = TRUE) %>% select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp = Temp...C..LGR.S.N..21435200..SEN.S.N..21435200.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%r"))

tank3a <- read.csv("../data/HOBO/yrv_5.csv", skip = 1, header = TRUE) %>% select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp = Temp...C..LGR.S.N..21435403..SEN.S.N..21435403.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%r"))
tank3b <- read.csv("../data/HOBO/yrv_6.csv", skip = 1, header = TRUE) %>% select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp = Temp...C..LGR.S.N..21435199..SEN.S.N..21435199.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%r"))

tank4a <- read.csv("../data/HOBO/yrv_7.csv", skip = 1, header = TRUE) %>% select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp = Temp...C..LGR.S.N..21435407..SEN.S.N..21435407.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%r"))
tank4b <- read.csv("../data/HOBO/yrv_8.csv", skip = 1, header = TRUE) %>% select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp = Temp...C..LGR.S.N..21435413..SEN.S.N..21435413.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%r"))

tank5a <- read.csv("../data/HOBO/yrv_9_13_comb.csv", skip = 1,  header = TRUE) %>% select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp = Temp...C..LGR.S.N..21435414..SEN.S.N..21435414.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%r"))
tank5b <- read.csv("../data/HOBO/yrv_10.csv", skip = 1, header = TRUE) %>% select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp = Temp...C..LGR.S.N..21435411..SEN.S.N..21435411.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%r"))

tank6a <- read.csv("../data/HOBO/yrv_11.csv", skip = 1, header = TRUE) %>% select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp = Temp...C..LGR.S.N..21435402..SEN.S.N..21435402.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%r"))
tank6b <- read.csv("../data/HOBO/yrv_12.csv", skip = 1, header = TRUE) %>% select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp = Temp...C..LGR.S.N..21435410..SEN.S.N..21435410.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%r"))
```

```{r}
head(tank1a) #Confirm import and formatting
```

## Treatment conditions

### Calculate treatment mean and SE

Row 631 is when HOBO loggers are recording data on the first full day of the experiment after the temperatures have reached the treatment setpoints, so I'll use that as my cutoff.

```{r}
treatmentTempStats <- data.frame("tank" = seq(1:6),
                                 "meanTemp" = rep(0, times = 6),
                                 "sdTemp" = rep(0, times = 6)) #Create an empty dataframe
```

```{r}
hobos <- list(tank1a, tank2a, tank3a, tank4a, tank5a, tank6a,
              tank1b, tank2b, tank3b, tank4b, tank5b, tank6b) #Create a list of filenames to cycle through
```


```{r}
for (i in 1:(length(hobos)/2)) {
  treatmentTempStats[i,2] <- mean(rbind(hobos[[i]][,2][631:length(hobos[[i]][,2])], hobos[[i+6]][,2][631:length(hobos[[i+6]][,2])]), na.rm = TRUE)
  treatmentTempStats[i,3] <- sd(rbind(hobos[[i]][,2][631:length(hobos[[i]][,2])], hobos[[i+6]][,2][631:length(hobos[[i+6]][,2])]), na.rm = TRUE)
} #For each tank, combine the treatment temperature data from both HOBO loggers,  either find the average or SD, and save in treatmentTempStats
head(treatmentTempStats) #Confirm calculations
```










