---
title: "03-TTR-analysis"
author: "Yaamini Venkataraman"
date: '2023-04-10'
output: html_document
---

In this document I'll examine how time-to-right (TTR) varies by temperature and time.

# Set up R Markdown document

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = normalizePath("../output/03-TTR-analysis/")) #Set root directory
```

```{r}
getwd()
```

# Install packages

```{r packages, include=FALSE}
# install.packages("tidyverse")
# install.packages("plotrix")
# install.packages("RColorBrewer")
# install.packages("cowplot")
# install.packages("lme4")
install.packages("glmmTMB", type = "source")
# install.packages("DHARMa")
# install.packages("car")
# install.packages("emmeans")
# install.packages("multcompView")
# install.packages("multcomp")
# install.packages("patchwork")
library(tidyverse)
library(plotrix)
library(RColorBrewer)
library(cowplot)
library(lme4)
library(glmmTMB)
library(DHARMa)
library(car)
library(emmeans)
library(multcompView)
library(multcomp)
library(patchwork)
```

```{r}
sessionInfo()
```

# Import data

```{r}
rawTTR <- read.csv("../../data/time-to-right.csv", header = TRUE) #Import raw data
head(rawTTR) #Confirm import. Trial data is in seconds.
```

# Format data

```{r}
modTTR <- rawTTR %>%
  dplyr::select(., -c(sampling.ID, notes, probe.number)) %>% 
  filter(., number.swimmers != 0) %>%
  mutate(., day = case_when(date == "7/8/22" ~ 4,
                            date == "7/12/22" ~ 8,
                            date == "7/15/22" ~ 11,
                            date == "7/19/22" ~ 15,
                            date == "7/22/22" ~ 18,
                            date == "7/26/22" ~ 22)) %>%
  mutate(., treatment = case_when(treatment.tank == "1" | treatment.tank == "4" ~ "13C",
                                  treatment.tank == "2" | treatment.tank == "5" ~ "5C",
                                  treatment.tank == "3" | treatment.tank == "6" ~ "30C")) %>%
  mutate(., missing.legs = case_when(number.legs < 10 ~ "Y",
                                     number.legs == 10 ~ "N")) %>%
  mutate(., integument.cont = dplyr::recode(integument.color,
                                            "B" = 0,
                                            "BG" = 0.5,
                                            "G" = 1, 
                                            "YG" = 1.5, 
                                            "Y" = 2, 
                                            "YO" = 2.5 , 
                                            "O" = 3, 
                                            "RO" = 3.5)) %>%
  mutate(., trial.1 = na_if(trial.1, 91.00)) %>%
  mutate(., trial.2 = na_if(trial.2, 91.00)) %>%
  mutate(., trial.3 = na_if(trial.3, 91.00)) %>%
  rowwise(.) %>% 
  mutate(., TTRavg = mean(c(trial.1, trial.2, trial.3), na.rm = TRUE)) %>%
  filter(., is.na(TTRavg) == FALSE) %>%
  mutate(., TTRSE = std.error(c(trial.1, trial.2, trial.3), na.rm = TRUE)) %>%
  filter(., is.na(TTRSE) == FALSE) %>%
  group_by(., day, treatment) %>%
  mutate(., TTRavgFull = mean(TTRavg)) %>%
  mutate(., TTRSEFull = std.error(TTRavg)) %>%
  mutate(., TTRavgFullLow = TTRavgFull - TTRSEFull) %>%
  mutate(., TTRavgFullHigh = TTRavgFull + TTRSEFull) %>%
  ungroup(.) #Remove sampling ID and notes columns. Use slice to remove rows for crabs that were missing both swimmers. Add new column with day information. Add new column with treatment information. Create a a new column that with an index for width/length. Create new column as a binary for whether or not crabs are missing legs. Change all 91 to NA, calculate average and SE using data from the three TTR trials using rowwise operations, and remove rows where TTRavg | TTRSE = NA.  Calculate average and SE for all samples in a treatment for each day. Add/subtract TTRSEFull to/from TTRavgFull to get bounds. Ungroup.
head(modTTR) #Confirm formatting
```

# Data exploration

## Frequency of no flips

```{r}
rawTTR %>%
  dplyr::select(., -c(sampling.ID, notes, probe.number)) %>%
  mutate(., day = case_when(date == "7/8/22" ~ 4,
                            date == "7/12/22" ~ 8,
                            date == "7/15/22" ~ 11,
                            date == "7/19/22" ~ 15,
                            date == "7/22/22" ~ 18,
                            date == "7/26/22" ~ 22)) %>%
  mutate(., treatment = case_when(treatment.tank == "1" | treatment.tank == "4" ~ "13C",
                                  treatment.tank == "2" | treatment.tank == "5" ~ "5C",
                                  treatment.tank == "3" | treatment.tank == "6" ~ "30C")) %>%
  mutate(., missing.legs = case_when(number.legs < 10 ~ "Y",
                                     number.legs == 10 ~ "N")) %>%
  filter(., trial.1 == 91 | trial.2 == 91 | trial.3 == 91) #Take raw TTR data and select columns of interest. Create a day column and modify treatment column. Create a missing.legs column. Filter data where crabs did not right within 90 s for any trial
```

```{r}
rawTTR %>%
  dplyr::select(., -c(sampling.ID, notes, probe.number)) %>%
  mutate(., day = case_when(date == "7/8/22" ~ 4,
                            date == "7/12/22" ~ 8,
                            date == "7/15/22" ~ 11,
                            date == "7/19/22" ~ 15,
                            date == "7/22/22" ~ 18,
                            date == "7/26/22" ~ 22)) %>%
  mutate(., treatment = case_when(treatment.tank == "1" | treatment.tank == "4" ~ "13C",
                                  treatment.tank == "2" | treatment.tank == "5" ~ "5C",
                                  treatment.tank == "3" | treatment.tank == "6" ~ "30C")) %>%
  mutate(., missing.legs = case_when(number.legs < 10 ~ "Y",
                                     number.legs == 10 ~ "N")) %>%
  filter(., trial.1 == 91 | trial.2 == 91 | trial.3 == 91) %>%
  write.csv(., "TTR-no-flips.csv", quote = FALSE, row.names = FALSE) #Take raw TTR data and select columns of interest. Create a day column and modify treatment column. Create a missing.legs column. Filter data where crabs did not right within 90 s for any trial. Save as a CSV
```

## TTR by various factors

```{bash}
mkdir figures
```

```{r}
plotColors <- c(brewer.pal(9, "Reds")[7],
                brewer.pal(9, "Greys")[7],
                brewer.pal(9, "Blues")[7]) #Create color scheme
```

### Treatment

```{r}
modTTR %>%
  ggplot(., mapping = aes(x = day, y = TTRavg, color = treatment)) +
  geom_jitter(alpha = 0.75) +
  scale_x_continuous(name = "Day",
                     breaks = unique(modTTR$day),
                     limits = c(4, 22)) +
  scale_y_continuous(name = "Average Time-to-Right (s)",
                     breaks = seq(0, 50, 10),
                     limits = c(0, 50)) +
  scale_color_manual(values = c(plotColors[2], plotColors[1], plotColors[3]), 
                     name = "Temperature (ºC)",
                     breaks = c("13C", "30C", "5C"),
                     labels = c("13", "30", "5")) +
  theme_classic(base_size = 15) #Plot average TTR and add lm fits without SE bars. Modify x-axis to only show experimental days where TTR was measured. Scale y axis to include 91 = indication that crabs did not right within 90 s. Assign colors to each treatment. Increase base font size.
ggsave("figures/time-to-right-raw.pdf", height = 8.5, width = 11)
```

I'm now going to make plots for various other factors to see if any of these need to be taken into consideration!

### Sex

```{r}
modTTR %>%
  ggplot(., mapping = aes(x = day, y = TTRavg, color = sex)) +
  geom_jitter(alpha = 0.75) +
  scale_x_continuous(name = "Day",
                     breaks = unique(modTTR$day),
                     limits = c(4, 22)) +
  scale_y_continuous(name = "Average Time-to-Right (s)",
                     breaks = seq(0, 50, 10),
                     limits = c(0, 50)) +
  theme_classic(base_size = 15)
ggsave("figures/time-to-right-sex.pdf", height = 8.5, width = 11)
```

### Tank

```{r}
modTTR %>%
  ggplot(., mapping = aes(x = day, y = TTRavg, color = as.factor(treatment.tank), shape = as.factor(treatment.tank))) +
  geom_jitter(alpha = 0.75) +
  scale_x_continuous(name = "Day",
                     breaks = unique(modTTR$day),
                     limits = c(4, 22)) +
  scale_y_continuous(name = "Average Time-to-Right (s)",
                     breaks = seq(0, 50, 10),
                     limits = c(0, 50)) +
  scale_color_manual(values = rep(c(plotColors[2], plotColors[3], plotColors[1]), times = 2),
                     name = "Tank",
                     breaks = c("1", "2", "3", "4", "5", "6")) +
  scale_shape_manual(values = c(19, 19, 19, 17, 17, 17),
                     name = "Tank",
                     breaks = c("1", "2", "3", "4", "5", "6")) +
  theme_classic(base_size = 15)
ggsave("figures/time-to-right-tank.pdf", height = 8.5, width = 11)
```

### Integument color

```{r}
modTTR %>%
  ggplot(., mapping = aes(x = day, y = TTRavg, color = integument.color)) +
  geom_jitter(alpha = 0.8) +
  scale_x_continuous(name = "Day",
                     breaks = unique(modTTR$day),
                     limits = c(4, 22)) +
  scale_y_continuous(name = "Average Time-to-Right (s)",
                     breaks = seq(0, 50, 10),
                     limits = c(0, 50)) +
  scale_color_manual(values = c("#66C2A5","#A6D96A","#D9EF8B","#FFFF99","#FDAE61","#F46D43","#D53E4F"),
                     name = "Integument",
                     breaks = c("BG", "G", "YG", "Y", "YO", "O", "RO"),
                     labels = c("BG", "G", "YG", "Y", "YO", "O", "RO")) +
  theme_classic(base_size = 15)
ggsave("figures/time-to-right-integument.pdf", height = 8.5, width = 11)
```

### Number of Legs

```{r}
modTTR %>%
  ggplot(., mapping = aes(x = day, y = TTRavg, color = as.factor(number.other.legs))) +
  geom_jitter(alpha = 0.8) +
  scale_x_continuous(name = "Day",
                     breaks = unique(modTTR$day),
                     limits = c(4, 22)) +
  scale_y_continuous(name = "Average Time-to-Right (s)",
                     breaks = seq(0, 50, 10),
                     limits = c(0, 50)) +
  theme_classic(base_size = 15)
ggsave("figures/time-to-right-numlegs.pdf", height = 8.5, width = 11)
```

### Number of Swimmers

```{r}
modTTR %>%
  ggplot(., mapping = aes(x = day, y = TTRavg, color = as.factor(number.swimmers))) +
  geom_jitter(alpha = 0.8) +
  scale_x_continuous(name = "Day",
                     breaks = unique(modTTR$day),
                     limits = c(4, 22)) +
  scale_y_continuous(name = "Average Time-to-Right (s)",
                     breaks = seq(0, 50, 10),
                     limits = c(0, 50)) +
  theme_classic(base_size = 15)
ggsave("figures/time-to-right-numSwim.pdf", height = 8.5, width = 11)
```

### Weight

```{r}
modTTR %>%
  ggplot(., mapping = aes(x = day, y = TTRavg, color = weight)) +
  geom_jitter(alpha = 0.8) +
  scale_x_continuous(name = "Day",
                     breaks = unique(modTTR$day),
                     limits = c(4, 22)) +
  scale_y_continuous(name = "Average Time-to-Right (s)",
                     breaks = seq(0, 50, 10),
                     limits = c(0, 50)) +
  theme_classic(base_size = 15)
ggsave("figures/time-to-right-weight.pdf", height = 8.5, width = 11)
```

### Carapace width

```{r}
modTTR %>%
  ggplot(., mapping = aes(x = day, y = TTRavg, color = carapace.width)) +
  geom_jitter(alpha = 0.8) +
  scale_x_continuous(name = "Day",
                     breaks = unique(modTTR$day),
                     limits = c(4, 22)) +
  scale_y_continuous(name = "Average Time-to-Right (s)",
                     breaks = seq(0, 50, 10),
                     limits = c(0, 50)) +
  theme_classic(base_size = 15)
ggsave("figures/time-to-right-size.pdf", height = 8.5, width = 11)
```

# Assess differences by treatment using a GLMM

I have righting response data for crabs across time. I will use a gamma GLMM with a log-link to analyze how temperature, experimental day, and their interaction impact righting response. However, the crabs were unlabelled, meaning that there were likely some crabs that were assayed multiple times across the experiment. To account for this in the data analysis, I will use tank as a random effect.

## GLMM

My first step is to try a few different model families with this dataset to understand which one leads to the lowest AIC.

```{r}
#Scale continuous variables for the GLMM

modTTR$carapace.width.sc <- scale(modTTR$carapace.width)
modTTR$day.sc <- scale(modTTR$day)
modTTR$integument.sc <- scale(modTTR$integument.cont)
modTTR$number.swimmers.sc <- scale(modTTR$number.swimmers)
```

### Gamma model family

```{r}
TTRmodelGamma <- glmmTMB( #Generalized linear mixed effect model
  TTRavg ~ as.factor(treatment) + day.sc + as.factor(treatment):day.sc + #Treatment, day, and their interaction
    as.factor(sex) + integument.sc + carapace.width.sc + number.swimmers.sc + #Demographic variables: sex, integument color, carapace width, and number of swimmers
    (1 | treatment.tank), #Use tank as a random effect to account for potential dependence between data
  data = modTTR,
  family = Gamma(link = "log") #Gamma model with a log link
)
```

```{r}
drop1(TTRmodelGamma) #Use a backwards deletion approach and AIC to identify best fit model. 

#Removing number.swimmers.sc leads to the lowest AIC
```
```{r}
TTRmodelGamma2 <- glmmTMB( #Generalized linear mixed effect model
  TTRavg ~ as.factor(treatment) + day.sc + as.factor(treatment):day.sc + #Treatment, day, and their interaction
    as.factor(sex) + integument.sc + carapace.width.sc + #Demographic variables: sex, integument color, carapace width
    (1 | treatment.tank), #Use tank as a random effect to account for potential dependence between data
  data = modTTR,
  family = Gamma(link = "log") #Gamma model with a log link
)
```

```{r}
drop1(TTRmodelGamma2) #Use a backwards deletion approach and AIC to identify best fit model. 

#Removing carapace.width.sc leads to the lowest AIC
```
```{r}
TTRmodelGamma3 <- glmmTMB( #Generalized linear mixed effect model
  TTRavg ~ as.factor(treatment) + day.sc + as.factor(treatment):day.sc + #Treatment, day, and their interaction
    as.factor(sex) + integument.sc + #Demographic variables: sex, integument color
    (1 | treatment.tank), #Use tank as a random effect to account for potential dependence between data
  data = modTTR,
  family = Gamma(link = "log") #Gamma model with a log link
)
```

```{r}
drop1(TTRmodelGamma3) #Use a backwards deletion approach and AIC to identify best fit model. 

#Removing integument.sc leads to the lowest AIC
```

```{r}
TTRmodelGamma4 <- glmmTMB( #Generalized linear mixed effect model
  TTRavg ~ as.factor(treatment) + day.sc + as.factor(treatment):day.sc + #Treatment, day, and their interaction
    as.factor(sex) + #Demographic variables: sex
    (1 | treatment.tank), #Use tank as a random effect to account for potential dependence between data
  data = modTTR,
  family = Gamma(link = "log") #Gamma model with a log link
)
```

```{r}
drop1(TTRmodelGamma4) #Use a backwards deletion approach and AIC to identify best fit model. 

#No more terms to delete!
```

### Gamma model family with polynomial term

```{r}
TTRmodelGammaPoly <- glmmTMB( #Generalized linear mixed effect model
  TTRavg ~ as.factor(treatment)*(day.sc + I(day.sc^2)) + #Treatment, day (single and polynomial), and their interaction
    as.factor(sex) + integument.sc + carapace.width.sc + number.swimmers.sc + #Demographic variables: sex, integument color, carapace width, and number of swimmers
    (1 | treatment.tank), #Use tank as a random effect to account for potential dependence between data
  data = modTTR,
  family = Gamma(link = "log") #Gamma model with a log link
)
```

```{r}
drop1(TTRmodelGammaPoly) #Use a backwards deletion approach and AIC to identify best fit model. 

#Dropping number.swimmers.sc term leads to the lowest AIC
```

```{r}
TTRmodelGammaPoly2 <- glmmTMB( #Generalized linear mixed effect model
  TTRavg ~ as.factor(treatment)*(day.sc + I(day.sc^2)) + #Treatment, day (single and polynomial), and their interaction
    as.factor(sex) + integument.sc + carapace.width.sc + #Demographic variables: sex, integument color, carapace width
    (1 | treatment.tank), #Use tank as a random effect to account for potential dependence between data
  data = modTTR,
  family = Gamma(link = "log") #Gamma model with a log link
)
```

```{r}
drop1(TTRmodelGammaPoly2) #Use a backwards deletion approach and AIC to identify best fit model. 

#Dropping carapace.width.sc term leads to the lowest AIC
```

```{r}
TTRmodelGammaPoly3 <- glmmTMB( #Generalized linear mixed effect model
  TTRavg ~ as.factor(treatment)*(day.sc + I(day.sc^2)) + #Treatment, day (single and polynomial), and their interaction
    as.factor(sex) + integument.sc + #Demographic variables: sex, integument color
    (1 | treatment.tank), #Use tank as a random effect to account for potential dependence between data
  data = modTTR,
  family = Gamma(link = "log") #Gamma model with a log link
)
```

```{r}
drop1(TTRmodelGammaPoly3) #Use a backwards deletion approach and AIC to identify best fit model. 

#Dropping integument.sc term leads to the lowest AIC
```

```{r}
TTRmodelGammaPoly4 <- glmmTMB( #Generalized linear mixed effect model
  TTRavg ~ as.factor(treatment)*(day.sc + I(day.sc^2)) + #Treatment, day (single and polynomial), and their interaction
    as.factor(sex) + #Demographic variables: sex
    (1 | treatment.tank), #Use tank as a random effect to account for potential dependence between data
  data = modTTR,
  family = Gamma(link = "log") #Gamma model with a log link
)
```

```{r}
drop1(TTRmodelGammaPoly4) #Use a backwards deletion approach and AIC to identify best fit model. 

#Dropping polynomial term leads to the lowest AIC
```

Since dropping the polynomial term will lead to the lowest AIC, I will not go through the model selection process anymore, as this makes the model identical to the previous gamma family model without the polynomial term.

### Log normal model family

```{r}
TTRmodelLogNorm <- glmmTMB( #Generalized linear mixed effect model
  TTRavg ~ as.factor(treatment) + day.sc + as.factor(treatment):day.sc + #Treatment, day, and their interaction
    as.factor(sex) + integument.sc + carapace.width.sc + number.swimmers.sc + #Demographic variables: sex, integument color, carapace width, and number of swimmers
    (1 | treatment.tank), #Use tank as a random effect to account for potential dependence between data
  data = modTTR,
  family = lognormal(link = "log") #Log normal model with a log link
)
```

```{r}
drop1(TTRmodelLogNorm) #Use a backwards deletion approach and AIC to identify best fit model. 

#Dropping the interaction term leads to the lowest AIC
```

```{r}
#Rerun the model without temperature:day interaction

TTRmodelLogNorm2 <- glmmTMB( #Generalized linear mixed effect model
  TTRavg ~ as.factor(treatment) + day.sc + #Treatment and day
    as.factor(sex) + integument.sc + carapace.width.sc + number.swimmers.sc + #Demographic variables: sex, integument color, carapace width, number of swimmers
    (1 | treatment.tank), #Use tank as a random effect to account for potential dependence between data
  data = modTTR,
  family = lognormal(link = "log") #Log normal model with a log link
)
```

```{r}
drop1(TTRmodelLogNorm2) #Use a backwards deletion approach and AIC to identify best fit model. 

#Dropping integument.sc will reduce AIC the most
```

```{r}
#Rerun the model without integument.sc

TTRmodelLogNorm3 <- glmmTMB( #Generalized linear mixed effect model
  TTRavg ~ as.factor(treatment) + day.sc + #Treatment and day
    as.factor(sex) + carapace.width.sc + number.swimmers.sc + #Demographic variables: sex, carapace width, number of swimmers
    (1 | treatment.tank), #Use tank as a random effect to account for potential dependence between data
  data = modTTR,
  family = lognormal(link = "log") #Log normal model with a log link
)
```

```{r}
drop1(TTRmodelLogNorm3) #Use a backwards deletion approach and AIC to identify best fit model. 

#Dropping day.sc will reduce AIC the most
```

```{r}
TTRmodelLogNorm4 <- glmmTMB( #Generalized linear mixed effect model
  TTRavg ~ as.factor(treatment) + #Treatment
    as.factor(sex) + carapace.width.sc + number.swimmers.sc + #Demographic variables: sex, carapace width, number of swimmers
    (1 | treatment.tank), #Use tank as a random effect to account for potential dependence between data
  data = modTTR,
  family = lognormal(link = "log") #Log normal model with a log link
)
```

```{r}
drop1(TTRmodelLogNorm4) #Use a backwards deletion approach and AIC to identify best fit model. 

#Removing sex will reduce AIC the most
```

```{r}
TTRmodelLogNorm5 <- glmmTMB( #Generalized linear mixed effect model
  TTRavg ~ as.factor(treatment) + #Treatment
    carapace.width.sc + number.swimmers.sc + #Demographic variables: carapace width, number of swimmers
    (1 | treatment.tank), #Use tank as a random effect to account for potential dependence between data
  data = modTTR,
  family = lognormal(link = "log") #Log normal model with a log link
)
```

```{r}
drop1(TTRmodelLogNorm5) #Use a backwards deletion approach and AIC to identify best fit model. 

#No more terms to remove!
```

### Inverse gaussian model family

```{r}
TTRmodelInvGaus <- glmmTMB( #Generalized linear mixed effect model
  TTRavg ~ as.factor(treatment) + day.sc + as.factor(treatment):day.sc + #Treatment, day, and their interaction
    as.factor(sex) + integument.sc + carapace.width.sc + number.swimmers.sc + #Demographic variables: sex, integument color, carapace width, and number of swimmers
    (1 | treatment.tank), #Use tank as a random effect to account for potential dependence between data
  data = modTTR,
  family = inverse.gaussian(link = "log") #Log normal model with a log link
)
```

```{r}
drop1(TTRmodelInvGaus) #Use a backwards deletion approach and AIC to identify best fit model. 

#Removing the interaction term leads to the lowest AIC
```

```{r}
TTRmodelInvGaus2 <- glmmTMB( #Generalized linear mixed effect model
  TTRavg ~ as.factor(treatment) + day.sc +  #Treatment, day
    as.factor(sex) + integument.sc + carapace.width.sc + number.swimmers.sc + #Demographic variables: sex, integument color, carapace width, and number of swimmers
    (1 | treatment.tank), #Use tank as a random effect to account for potential dependence between data
  data = modTTR,
  family = inverse.gaussian(link = "log") #Log normal model with a log link
)
```

```{r}
drop1(TTRmodelInvGaus2) #Use a backwards deletion approach and AIC to identify best fit model. 

#Removing sex leads to the lowest AIC
```

```{r}
TTRmodelInvGaus3 <- glmmTMB( #Generalized linear mixed effect model
  TTRavg ~ as.factor(treatment) + day.sc +  #Treatment, day
    integument.sc + carapace.width.sc + number.swimmers.sc + #Demographic variables: integument color, carapace width, and number of swimmers
    (1 | treatment.tank), #Use tank as a random effect to account for potential dependence between data
  data = modTTR,
  family = inverse.gaussian(link = "log") #Log normal model with a log link
)
```

```{r}
drop1(TTRmodelInvGaus3) #Use a backwards deletion approach and AIC to identify best fit model. 

#Removing number.swimmers.sc leads to the lowest AIC
```

```{r}
TTRmodelInvGaus4 <- glmmTMB( #Generalized linear mixed effect model
  TTRavg ~ as.factor(treatment) + day.sc +  #Treatment, day
    integument.sc + carapace.width.sc + #Demographic variables: integument color, carapace width
    (1 | treatment.tank), #Use tank as a random effect to account for potential dependence between data
  data = modTTR,
  family = inverse.gaussian(link = "log") #Log normal model with a log link
)
```

```{r}
drop1(TTRmodelInvGaus4) #Use a backwards deletion approach and AIC to identify best fit model. 

#No more terms to delete!
```

## Final model

```{r}
AIC(TTRmodelGamma4, TTRmodelLogNorm5, TTRmodelInvGaus4) #Compare AIC amongst the most parsimonious models for different model families
```

Since the gamma model has the lowest AIC, I will use that model as the final model.

```{r}
summary(TTRmodelGamma4) #Summary output of the final model
```

## Check assumptions

```{r}
vif(TTRmodelGamma4) # Check multicollinearity assumption.

#Values < 5 are ideal, assumption is met
```

```{r}
sim_residuals <- simulateResiduals(fittedModel = TTRmodelGamma4, n = 1000, plot = TRUE) #Simulate residuals for the final model

#Residual plot is characteristic of underdispersion according to the DHARMa manual.
```

```{r}
testUniformity(sim_residuals) #Kolmogorov-Smirnov test
```

```{r}
testDispersion(sim_residuals) #Test dispersion

#Dispersion = 1.7266, but not significant overdispersion
```

```{r}
testOutliers(sim_residuals) #Test for outliers

#No significant outliers
```


# 3. Simulate Residuals with DHARMa

# The plot function output gives:
# - Q-Q plot for residual uniformity & distribution fit
# - Residuals vs. Fitted plot to test for heteroscedasticity and non-linearity

# 5. Check Normality of Random Effects (Tank)
ranef_tanks <- ranef(fit_gamma)$Tank$`(Intercept)`
qqnorm(ranef_tanks)
qqline(ranef_tanks)

## Save model output

```{r}
write.csv(broom::tidy(TTRmodel2), "TTR-glm-output.csv", quote = FALSE, row.names = FALSE) #Save table version of glm output
```

## Posthoc tests for significant predictors

There was a significant impacts of treatment! I will investigate these effects using post-hoc tests with `emmeans`.

```{r}
TTRmodel_treatment <- emmeans(TTRmodel2, pairwise ~ treatment, adjust = "tukey", data = modTTR) #Use emmeans to obtain pairwise treatment differences
TTRmodel_treatment$contrasts %>% broom::tidy(.) #Extract the contrast information to look at pairwise test output. Clean output using broom.
TTRmodel_treatment$contrasts %>% 
  broom::tidy(.) %>%
  write.csv("tukey-test-treatment.csv", quote = FALSE, row.names = FALSE) #Save output as a csv.
```

# Threshold analysis

Another way to analyze this data is to count how many crabs within each treatment righted within a certain threshold. I can identify this threshold by taking the upper bound of a boxplot, then using a binomial model similar to [Coyle et al. (2019)](https://doi.org/10.1242/jeb.203521).

## Define threshold

```{r}
boxplot.stats(modTTR$TTRavg)$stats[5] #Use Tukey's method to identify upper bound of outliers
```

```{r}
modTTR <- modTTR %>%
  mutate(., TTRthresh = case_when(TTRavg > boxplot.stats(modTTR$TTRavg)$stats[5] ~ 0,
                                  TTRavg <= boxplot.stats(modTTR$TTRavg)$stats[5] ~ 1)) #Create a new column where 1 = righted within the upper bound of the boxplot, 0 = failure to right within that threshold
head(modTTR) #Confirm changes
```

## Binomial model

```{r}
TTRthreshModel <- glm(I(TTRthresh == 1) ~ treatment + day + treatment*day + as.factor(sex) + as.factor(integument.color) + carapace.width + weight + as.factor(missing.legs), 
                      family = binomial(), 
                      data = modTTR) #Create a binomial model, where 1 = success of righting wihtin the defined threshold
summary(TTRthreshModel) #Only the contrast bewteen 5C and 13C control is significant
```

```{r}
step(TTRthreshModel, test = "LRT") #Use likelihood ratio tests and backwards deletion approaches to identify the best model
```

## Final model

```{r}
TTRthreshModel2 <- glm(I(TTRthresh == 1) ~ treatment + day,
                       family = binomial(),
                       data = modTTR) #Run model selected by step with only significant terms
summary(TTRthreshModel2) #control vs. 5C significantly different, significant impact of day
```

## Save model output

```{r}
write.csv(broom::tidy(TTRthreshModel2), "TTR-binomialModel.csv", quote = FALSE, row.names = FALSE) #Save binomal GLM output
```

# Final plot

```{r}
modTTR %>%
  dplyr::select(., c(day, treatment, TTRavgFull, TTRavgFullLow, TTRavgFullHigh)) %>%
  distinct(.) %>%
  ggplot(., mapping = aes(x = day, y = TTRavgFull, color = treatment, shape = treatment)) +
  geom_pointrange(aes(ymin = TTRavgFullLow,
                      ymax = TTRavgFullHigh), 
                  size = 0.5, alpha = 0.75) +
  geom_line(y = boxplot.stats(modTTR$TTRavg)$stats[5], lty = 2, color = "black") +
  scale_x_continuous(name = "Day",
                     breaks = unique(modTTR$day),
                     limits = c(4, 22)) +
  scale_y_continuous(name = "Average Time-to-Right (s)",
                     breaks = c(0, seq(1, 5, 2), seq(5, 20, 5)),
                     limits = c(0, 17)) +
  scale_color_manual(values = c(plotColors[2], plotColors[1], plotColors[3]), 
                     name = "Temperature (ºC)",
                     breaks = c("13C", "30C", "5C"),
                     labels = c("13", "30", "5")) +
  scale_shape_manual(values = c(19, 17, 15),
                     name = "Temperature (ºC)",
                     breaks = c("13C", "30C", "5C"),
                     labels = c("13", "30", "5")) +
  theme_classic(base_size = 15) #Plot average TTR and add lm fits without SE bars. Add vertical line for boxplot outlier threshold. Modify x-axis to only show experimental days where TTR was measured. Scale y axis to include 91 = indication that crabs did not right within 90 s. Assign colors to each treatment. Increase base font size.
#ggsave("figures/time-to-right-avg.pdf", height = 8.5, width = 11)
```

```{r}
modTTR %>%
  dplyr::select(., c(day, treatment, TTRavg)) %>%
  distinct(.) %>%
  mutate(., treatment = gsub(x = treatment, pattern = "5C", replacement = "05C")) %>%
  mutate(., day = gsub(x = day, pattern = 4, replacement = "04")) %>%
  mutate(., day = gsub(x = day, pattern = 8, replacement = "08")) %>%
  mutate(., day = gsub(x = day, pattern = 11, replacement = "11")) %>%
  mutate(., day = gsub(x = day, pattern = 108, replacement = "18")) %>%
  mutate(., day = gsub(x = day, pattern = 15, replacement = "15")) %>%
  mutate(., day = gsub(x = day, pattern = 22, replacement = "22")) %>%
  ggplot(mapping = aes(x = day, y = TTRavg, color = treatment, shape = treatment)) +
  geom_boxplot(outlier.shape = NA) + geom_jitter(position = position_dodge(width = 0.75)) +
  # geom_hline(aes(yintercept = boxplot.stats(modTTR$TTRavg)$stats[5]), lty = 2, color = "black") +
  ylab("Average Time-to-Right (s)") +
  scale_x_discrete(name = "Day",
                   breaks = c("04", "08", "11", "18", "15", "22"),
                   labels = c("4", "8", "11", "18", "15", "22")) +
  scale_color_manual(values = c(plotColors[3], plotColors[2], plotColors[1]), 
                     name = "Temperature (ºC)",
                     breaks = c("05C", "13C", "30C"),
                     labels = c("5", "13", "30")) +
  scale_shape_manual(values = c(15, 19, 17),
                     name = "Temperature (ºC)",
                     breaks = c("05C", "13C", "30C"),
                     labels = c("5", "13", "30")) +
  theme_classic(base_size = 15) #Select unique day, treatment, and TTRavg data. Replace numbers with characters to facilitate better ordering. Plot average TTR for each crab in a boxplot. Do not show outliers with the OG boxplot, but add them in with geom_jitter. Add horizontal line for boxplot outlier threshold. Assign colors to each treatment. Increase base font size.
#ggsave("figures/time-to-right-avg-boxplot.pdf", height = 8.5, width = 11)
```

```{r}
TTRmodel_treatment$contrasts %>% broom::tidy(.) #Extract the contrast information to look at pairwise test output. Clean output using broom.
```


```{r}
mainPlot <- modTTR %>%
  dplyr::select(., c(day, treatment, TTRavg)) %>%
  distinct(.) %>%
  mutate(., treatment = gsub(x = treatment, pattern = "5C", replacement = "05C")) %>%
  mutate(., day = gsub(x = day, pattern = 4, replacement = "04")) %>%
  mutate(., day = gsub(x = day, pattern = 8, replacement = "08")) %>%
  mutate(., day = gsub(x = day, pattern = 11, replacement = "11")) %>%
  mutate(., day = gsub(x = day, pattern = 108, replacement = "18")) %>%
  mutate(., day = gsub(x = day, pattern = 15, replacement = "15")) %>%
  mutate(., day = gsub(x = day, pattern = 22, replacement = "22")) %>%
  ggplot(mapping = aes(x = day, y = TTRavg, color = treatment, shape = treatment)) +
  geom_boxplot(outlier.shape = NA) + geom_jitter(position = position_dodge(width = 0.75)) +
  ylab("Average Time-to-Right (s)") +
  scale_x_discrete(name = "Day",
                   breaks = c("04", "08", "11", "18", "15", "22"),
                   labels = c("4", "8", "11", "18", "15", "22")) +
  scale_color_manual(values = c(plotColors[3], plotColors[2], plotColors[1]), 
                     name = "Temperature (ºC)",
                     breaks = c("05C", "13C", "30C"),
                     labels = c("5", "13", "30")) +
  scale_shape_manual(values = c(15, 19, 17),
                     name = "Temperature (ºC)",
                     breaks = c("05C", "13C", "30C"),
                     labels = c("5", "13", "30")) +
  annotate(geom = "text", x = 4.5, y = 32,
           label = expression(paste(bolditalic("13ºC vs. 30ºC: "), "t = 2.66, p = 0.02")),
           hjust = 0) +
  annotate(geom = "text", x = 4.5, y = 30.5,
           label = expression(paste(bolditalic("13ºC vs. 5ºC: "), "t = -15.48, p < 0.0001")),
           hjust = 0) +
  annotate(geom = "text", x = 4.5, y = 29,
           label = expression(paste(bolditalic("Day: "), "t = -1.55, p = 0.12")),
           hjust = 0) +
  annotate(geom = "text", x = 4.5, y = 27.5,
           label = expression(paste(bolditalic("Crab weight: "), "t = 1.47, p = 0.14")),
           hjust = 0) +
  theme_classic(base_size = 15) #Select unique day, treatment, and TTRavg data. Replace numbers with characters to facilitate better ordering. Plot average TTR for each crab in a boxplot. Do not show outliers with the OG boxplot, but add them in with geom_jitter. Assign colors to each treatment. Increase base font size.

insetPlot <-  modTTR %>%
  dplyr::select(., c(day, treatment, TTRavg)) %>%
  distinct(.) %>%
  filter(., treatment != "5C") %>%
  mutate(., day = gsub(x = day, pattern = 4, replacement = "04")) %>%
  mutate(., day = gsub(x = day, pattern = 8, replacement = "08")) %>%
  mutate(., day = gsub(x = day, pattern = 11, replacement = "11")) %>%
  mutate(., day = gsub(x = day, pattern = 108, replacement = "18")) %>%
  mutate(., day = gsub(x = day, pattern = 15, replacement = "15")) %>%
  mutate(., day = gsub(x = day, pattern = 22, replacement = "22")) %>%
  ggplot(mapping = aes(x = day, y = TTRavg, color = treatment, shape = treatment)) +
  geom_boxplot(outlier.shape = NA) + geom_jitter(position = position_dodge(width = 0.75)) +
  ylab("") +
  scale_x_discrete(name = "",
                   breaks = c("04", "08", "11", "18", "15", "22"),
                   labels = c("4", "8", "11", "18", "15", "22")) +
  scale_color_manual(values = c(plotColors[2], plotColors[1]), 
                     name = "Temperature (ºC)",
                     breaks = c("13C", "30C"),
                     labels = c("13", "30")) +
  scale_shape_manual(values = c(19, 17),
                     name = "Temperature (ºC)",
                     breaks = c("13C", "30C"),
                     labels = c("13", "30")) +
  scale_y_continuous(limits = c(0,3)) +
  theme_classic() + theme(legend.position = "none",
                          axis.line.x.bottom = element_line(linewidth = 0),
                          axis.line.y.left = element_line(linewidth = 0),
                          panel.border = element_rect(colour = "black", fill = NA, size = 1))

plotWithInset <- ggdraw() +
  draw_plot(mainPlot) +
  draw_plot(insetPlot, x = 0.3, y = 0.73, width = 0.52, height = 0.25)

plotWithInset

ggsave("figures/time-to-right-avg-boxplot-noThresh.pdf", height = 8.5, width = 11)
```
