---
title: "03-TTR-analysis"
author: "Yaamini Venkataraman"
date: '2023-04-10'
output: html_document
---

In this document I'll examine how time-to-right (TTR) varies by temperature and time.

# Set up R Markdown document

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = normalizePath("../output/03-TTR-analysis/")) #Set root directory
```

```{r}
getwd()
```

# Install packages

```{r packages, include=FALSE}
# install.packages("tidyverse")
# install.packages("plotrix")
# install.packages("RColorBrewer")
# install.packages("cowplot")
# install.packages("lme4")
# install.packages("glmmTMB", type = "source")
# install.packages("DHARMa")
# install.packages("performance")
# install.packages("see")
# install.packages("car")
# install.packages("parameters")
# install.packages("emmeans")
# install.packages("patchwork")
library(tidyverse)
library(plotrix)
library(RColorBrewer)
library(cowplot)
library(lme4)
library(glmmTMB)
library(DHARMa)
library(performance)
library(see)
library(car)
library(parameters)
library(emmeans)
library(patchwork)
```

```{r}
sessionInfo()
```

# Import data

```{r}
rawTTR <- read.csv("../../data/time-to-right.csv", header = TRUE) #Import raw data
head(rawTTR) #Confirm import. Trial data is in seconds.
```

# Format data

```{r}
modTTR <- rawTTR %>%
  dplyr::select(., -c(sampling.ID, notes, probe.number)) %>% 
  filter(., number.swimmers != 0) %>%
  mutate(., day = case_when(date == "7/8/22" ~ 4,
                            date == "7/12/22" ~ 8,
                            date == "7/15/22" ~ 11,
                            date == "7/19/22" ~ 15,
                            date == "7/22/22" ~ 18,
                            date == "7/26/22" ~ 22)) %>%
  mutate(., treatment = case_when(treatment.tank == "1" | treatment.tank == "4" ~ "13C",
                                  treatment.tank == "2" | treatment.tank == "5" ~ "5C",
                                  treatment.tank == "3" | treatment.tank == "6" ~ "30C")) %>%
  mutate(., missing.legs = case_when(number.legs < 10 ~ "Y",
                                     number.legs == 10 ~ "N")) %>%
  mutate(., integument.cont = dplyr::recode(integument.color,
                                            "B" = 0,
                                            "BG" = 0.5,
                                            "G" = 1, 
                                            "YG" = 1.5, 
                                            "Y" = 2, 
                                            "YO" = 2.5 , 
                                            "O" = 3, 
                                            "RO" = 3.5)) %>%
  mutate(., trial.1 = na_if(trial.1, 91.00)) %>%
  mutate(., trial.2 = na_if(trial.2, 91.00)) %>%
  mutate(., trial.3 = na_if(trial.3, 91.00)) %>%
  rowwise(.) %>% 
  mutate(., TTRavg = mean(c(trial.1, trial.2, trial.3), na.rm = TRUE)) %>%
  filter(., is.na(TTRavg) == FALSE) %>%
  mutate(., TTRSE = std.error(c(trial.1, trial.2, trial.3), na.rm = TRUE)) %>%
  filter(., is.na(TTRSE) == FALSE) %>%
  group_by(., day, treatment) %>%
  mutate(., TTRavgFull = mean(TTRavg)) %>%
  mutate(., TTRSEFull = std.error(TTRavg)) %>%
  mutate(., TTRavgFullLow = TTRavgFull - TTRSEFull) %>%
  mutate(., TTRavgFullHigh = TTRavgFull + TTRSEFull) %>%
  ungroup(.) #Remove sampling ID and notes columns. Use slice to remove rows for crabs that were missing both swimmers. Add new column with day information. Add new column with treatment information. Create a a new column that with an index for width/length. Create new column as a binary for whether or not crabs are missing legs. Change all 91 to NA, calculate average and SE using data from the three TTR trials using rowwise operations, and remove rows where TTRavg | TTRSE = NA.  Calculate average and SE for all samples in a treatment for each day. Add/subtract TTRSEFull to/from TTRavgFull to get bounds. Ungroup.
head(modTTR) #Confirm formatting
```

# Data exploration

## Frequency of no flips

```{r}
rawTTR %>%
  dplyr::select(., -c(sampling.ID, notes, probe.number)) %>%
  mutate(., day = case_when(date == "7/8/22" ~ 4,
                            date == "7/12/22" ~ 8,
                            date == "7/15/22" ~ 11,
                            date == "7/19/22" ~ 15,
                            date == "7/22/22" ~ 18,
                            date == "7/26/22" ~ 22)) %>%
  mutate(., treatment = case_when(treatment.tank == "1" | treatment.tank == "4" ~ "13C",
                                  treatment.tank == "2" | treatment.tank == "5" ~ "5C",
                                  treatment.tank == "3" | treatment.tank == "6" ~ "30C")) %>%
  mutate(., missing.legs = case_when(number.legs < 10 ~ "Y",
                                     number.legs == 10 ~ "N")) %>%
  filter(., trial.1 == 91 | trial.2 == 91 | trial.3 == 91) #Take raw TTR data and select columns of interest. Create a day column and modify treatment column. Create a missing.legs column. Filter data where crabs did not right within 90 s for any trial
```

```{r}
rawTTR %>%
  dplyr::select(., -c(sampling.ID, notes, probe.number)) %>%
  mutate(., day = case_when(date == "7/8/22" ~ 4,
                            date == "7/12/22" ~ 8,
                            date == "7/15/22" ~ 11,
                            date == "7/19/22" ~ 15,
                            date == "7/22/22" ~ 18,
                            date == "7/26/22" ~ 22)) %>%
  mutate(., treatment = case_when(treatment.tank == "1" | treatment.tank == "4" ~ "13C",
                                  treatment.tank == "2" | treatment.tank == "5" ~ "5C",
                                  treatment.tank == "3" | treatment.tank == "6" ~ "30C")) %>%
  mutate(., missing.legs = case_when(number.legs < 10 ~ "Y",
                                     number.legs == 10 ~ "N")) %>%
  filter(., trial.1 == 91 | trial.2 == 91 | trial.3 == 91) %>%
  write.csv(., "TTR-no-flips.csv", quote = FALSE, row.names = FALSE) #Take raw TTR data and select columns of interest. Create a day column and modify treatment column. Create a missing.legs column. Filter data where crabs did not right within 90 s for any trial. Save as a CSV
```

## TTR by various factors

```{bash}
mkdir figures
```

```{r}
plotColors <- c(brewer.pal(9, "Reds")[7],
                brewer.pal(9, "Greys")[7],
                brewer.pal(9, "Blues")[7]) #Create color scheme
```

### Treatment

```{r}
modTTR %>%
  ggplot(., mapping = aes(x = day, y = TTRavg, color = treatment)) +
  geom_jitter(alpha = 0.75) +
  scale_x_continuous(name = "Day",
                     breaks = unique(modTTR$day),
                     limits = c(4, 22)) +
  scale_y_continuous(name = "Average Time-to-Right (s)",
                     breaks = seq(0, 50, 10),
                     limits = c(0, 50)) +
  scale_color_manual(values = c(plotColors[2], plotColors[1], plotColors[3]), 
                     name = "Temperature (ºC)",
                     breaks = c("13C", "30C", "5C"),
                     labels = c("13", "30", "5")) +
  theme_classic(base_size = 15) #Plot average TTR and add lm fits without SE bars. Modify x-axis to only show experimental days where TTR was measured. Scale y axis to include 91 = indication that crabs did not right within 90 s. Assign colors to each treatment. Increase base font size.
ggsave("figures/time-to-right-raw.pdf", height = 8.5, width = 11)
```

I'm now going to make plots for various other factors to see if any of these need to be taken into consideration!

### Sex

```{r}
modTTR %>%
  ggplot(., mapping = aes(x = day, y = TTRavg, color = sex)) +
  geom_jitter(alpha = 0.75) +
  scale_x_continuous(name = "Day",
                     breaks = unique(modTTR$day),
                     limits = c(4, 22)) +
  scale_y_continuous(name = "Average Time-to-Right (s)",
                     breaks = seq(0, 50, 10),
                     limits = c(0, 50)) +
  theme_classic(base_size = 15)
ggsave("figures/time-to-right-sex.pdf", height = 8.5, width = 11)
```

### Tank

```{r}
modTTR %>%
  ggplot(., mapping = aes(x = day, y = TTRavg, color = as.factor(treatment.tank), shape = as.factor(treatment.tank))) +
  geom_jitter(alpha = 0.75) +
  scale_x_continuous(name = "Day",
                     breaks = unique(modTTR$day),
                     limits = c(4, 22)) +
  scale_y_continuous(name = "Average Time-to-Right (s)",
                     breaks = seq(0, 50, 10),
                     limits = c(0, 50)) +
  scale_color_manual(values = rep(c(plotColors[2], plotColors[3], plotColors[1]), times = 2),
                     name = "Tank",
                     breaks = c("1", "2", "3", "4", "5", "6")) +
  scale_shape_manual(values = c(19, 19, 19, 17, 17, 17),
                     name = "Tank",
                     breaks = c("1", "2", "3", "4", "5", "6")) +
  theme_classic(base_size = 15)
ggsave("figures/time-to-right-tank.pdf", height = 8.5, width = 11)
```

### Integument color

```{r}
modTTR %>%
  ggplot(., mapping = aes(x = day, y = TTRavg, color = integument.color)) +
  geom_jitter(alpha = 0.8) +
  scale_x_continuous(name = "Day",
                     breaks = unique(modTTR$day),
                     limits = c(4, 22)) +
  scale_y_continuous(name = "Average Time-to-Right (s)",
                     breaks = seq(0, 50, 10),
                     limits = c(0, 50)) +
  scale_color_manual(values = c("#66C2A5","#A6D96A","#D9EF8B","#FFFF99","#FDAE61","#F46D43","#D53E4F"),
                     name = "Integument",
                     breaks = c("BG", "G", "YG", "Y", "YO", "O", "RO"),
                     labels = c("BG", "G", "YG", "Y", "YO", "O", "RO")) +
  theme_classic(base_size = 15)
ggsave("figures/time-to-right-integument.pdf", height = 8.5, width = 11)
```

### Number of Legs

```{r}
modTTR %>%
  ggplot(., mapping = aes(x = day, y = TTRavg, color = as.factor(number.other.legs))) +
  geom_jitter(alpha = 0.8) +
  scale_x_continuous(name = "Day",
                     breaks = unique(modTTR$day),
                     limits = c(4, 22)) +
  scale_y_continuous(name = "Average Time-to-Right (s)",
                     breaks = seq(0, 50, 10),
                     limits = c(0, 50)) +
  theme_classic(base_size = 15)
ggsave("figures/time-to-right-numlegs.pdf", height = 8.5, width = 11)
```

### Number of Swimmers

```{r}
modTTR %>%
  ggplot(., mapping = aes(x = day, y = TTRavg, color = as.factor(number.swimmers))) +
  geom_jitter(alpha = 0.8) +
  scale_x_continuous(name = "Day",
                     breaks = unique(modTTR$day),
                     limits = c(4, 22)) +
  scale_y_continuous(name = "Average Time-to-Right (s)",
                     breaks = seq(0, 50, 10),
                     limits = c(0, 50)) +
  theme_classic(base_size = 15)
ggsave("figures/time-to-right-numSwim.pdf", height = 8.5, width = 11)
```

### Weight

```{r}
modTTR %>%
  ggplot(., mapping = aes(x = day, y = TTRavg, color = weight)) +
  geom_jitter(alpha = 0.8) +
  scale_x_continuous(name = "Day",
                     breaks = unique(modTTR$day),
                     limits = c(4, 22)) +
  scale_y_continuous(name = "Average Time-to-Right (s)",
                     breaks = seq(0, 50, 10),
                     limits = c(0, 50)) +
  theme_classic(base_size = 15)
ggsave("figures/time-to-right-weight.pdf", height = 8.5, width = 11)
```

### Carapace width

```{r}
modTTR %>%
  ggplot(., mapping = aes(x = day, y = TTRavg, color = carapace.width)) +
  geom_jitter(alpha = 0.8) +
  scale_x_continuous(name = "Day",
                     breaks = unique(modTTR$day),
                     limits = c(4, 22)) +
  scale_y_continuous(name = "Average Time-to-Right (s)",
                     breaks = seq(0, 50, 10),
                     limits = c(0, 50)) +
  theme_classic(base_size = 15)
ggsave("figures/time-to-right-size.pdf", height = 8.5, width = 11)
```

# Assess differences by treatment using a GLMM

I have righting response data for crabs across time. I will use a gamma GLMM with a log-link to analyze how temperature, experimental day, and their interaction impact righting response. However, the crabs were unlabelled, meaning that there were likely some crabs that were assayed multiple times across the experiment. To account for this in the data analysis, I will use tank as a random effect.

## GLMM

My first step is to try a few different model families with this dataset to understand which one leads to the lowest AIC.

```{r}
#Scale continuous variables for the GLMM

modTTR$carapace.width.sc <- scale(modTTR$carapace.width)
modTTR$day.sc <- scale(modTTR$day)
modTTR$integument.sc <- scale(modTTR$integument.cont)
modTTR$number.swimmers.sc <- scale(modTTR$number.swimmers)
```

### Gamma model family

```{r}
TTRmodelGamma <- glmmTMB( #Generalized linear mixed effect model
  TTRavg ~ as.factor(treatment) + day.sc + as.factor(treatment):day.sc + #Treatment, day, and their interaction
    as.factor(sex) + integument.sc + carapace.width.sc + number.swimmers.sc + #Demographic variables: sex, integument color, carapace width, and number of swimmers
    (1 | treatment.tank), #Use tank as a random effect to account for potential dependence between data
  data = modTTR,
  family = Gamma(link = "log") #Gamma model with a log link
)
```

```{r}
drop1(TTRmodelGamma) #Use a backwards deletion approach and AIC to identify best fit model. 

#Removing number.swimmers.sc leads to the lowest AIC
```
```{r}
TTRmodelGamma2 <- glmmTMB( #Generalized linear mixed effect model
  TTRavg ~ as.factor(treatment) + day.sc + as.factor(treatment):day.sc + #Treatment, day, and their interaction
    as.factor(sex) + integument.sc + carapace.width.sc + #Demographic variables: sex, integument color, carapace width
    (1 | treatment.tank), #Use tank as a random effect to account for potential dependence between data
  data = modTTR,
  family = Gamma(link = "log") #Gamma model with a log link
)
```

```{r}
drop1(TTRmodelGamma2) #Use a backwards deletion approach and AIC to identify best fit model. 

#Removing carapace.width.sc leads to the lowest AIC
```
```{r}
TTRmodelGamma3 <- glmmTMB( #Generalized linear mixed effect model
  TTRavg ~ as.factor(treatment) + day.sc + as.factor(treatment):day.sc + #Treatment, day, and their interaction
    as.factor(sex) + integument.sc + #Demographic variables: sex, integument color
    (1 | treatment.tank), #Use tank as a random effect to account for potential dependence between data
  data = modTTR,
  family = Gamma(link = "log") #Gamma model with a log link
)
```

```{r}
drop1(TTRmodelGamma3) #Use a backwards deletion approach and AIC to identify best fit model. 

#Removing integument.sc leads to the lowest AIC
```

```{r}
TTRmodelGamma4 <- glmmTMB( #Generalized linear mixed effect model
  TTRavg ~ as.factor(treatment) + day.sc + as.factor(treatment):day.sc + #Treatment, day, and their interaction
    as.factor(sex) + #Demographic variables: sex
    (1 | treatment.tank), #Use tank as a random effect to account for potential dependence between data
  data = modTTR,
  family = Gamma(link = "log") #Gamma model with a log link
)
```

```{r}
drop1(TTRmodelGamma4) #Use a backwards deletion approach and AIC to identify best fit model. 

#No more terms to delete!
```

### Gamma model family with polynomial term

```{r}
TTRmodelGammaPoly <- glmmTMB( #Generalized linear mixed effect model
  TTRavg ~ as.factor(treatment)*(day.sc + I(day.sc^2)) + #Treatment, day (single and polynomial), and their interaction
    as.factor(sex) + integument.sc + carapace.width.sc + number.swimmers.sc + #Demographic variables: sex, integument color, carapace width, and number of swimmers
    (1 | treatment.tank), #Use tank as a random effect to account for potential dependence between data
  data = modTTR,
  family = Gamma(link = "log") #Gamma model with a log link
)
```

```{r}
drop1(TTRmodelGammaPoly) #Use a backwards deletion approach and AIC to identify best fit model. 

#Dropping number.swimmers.sc term leads to the lowest AIC
```

```{r}
TTRmodelGammaPoly2 <- glmmTMB( #Generalized linear mixed effect model
  TTRavg ~ as.factor(treatment)*(day.sc + I(day.sc^2)) + #Treatment, day (single and polynomial), and their interaction
    as.factor(sex) + integument.sc + carapace.width.sc + #Demographic variables: sex, integument color, carapace width
    (1 | treatment.tank), #Use tank as a random effect to account for potential dependence between data
  data = modTTR,
  family = Gamma(link = "log") #Gamma model with a log link
)
```

```{r}
drop1(TTRmodelGammaPoly2) #Use a backwards deletion approach and AIC to identify best fit model. 

#Dropping carapace.width.sc term leads to the lowest AIC
```

```{r}
TTRmodelGammaPoly3 <- glmmTMB( #Generalized linear mixed effect model
  TTRavg ~ as.factor(treatment)*(day.sc + I(day.sc^2)) + #Treatment, day (single and polynomial), and their interaction
    as.factor(sex) + integument.sc + #Demographic variables: sex, integument color
    (1 | treatment.tank), #Use tank as a random effect to account for potential dependence between data
  data = modTTR,
  family = Gamma(link = "log") #Gamma model with a log link
)
```

```{r}
drop1(TTRmodelGammaPoly3) #Use a backwards deletion approach and AIC to identify best fit model. 

#Dropping integument.sc term leads to the lowest AIC
```

```{r}
TTRmodelGammaPoly4 <- glmmTMB( #Generalized linear mixed effect model
  TTRavg ~ as.factor(treatment)*(day.sc + I(day.sc^2)) + #Treatment, day (single and polynomial), and their interaction
    as.factor(sex) + #Demographic variables: sex
    (1 | treatment.tank), #Use tank as a random effect to account for potential dependence between data
  data = modTTR,
  family = Gamma(link = "log") #Gamma model with a log link
)
```

```{r}
drop1(TTRmodelGammaPoly4) #Use a backwards deletion approach and AIC to identify best fit model. 

#Dropping polynomial term leads to the lowest AIC
```

Since dropping the polynomial term will lead to the lowest AIC, I will not go through the model selection process anymore, as this makes the model identical to the previous gamma family model without the polynomial term.

### Log normal model family

```{r}
TTRmodelLogNorm <- glmmTMB( #Generalized linear mixed effect model
  TTRavg ~ as.factor(treatment) + day.sc + as.factor(treatment):day.sc + #Treatment, day, and their interaction
    as.factor(sex) + integument.sc + carapace.width.sc + number.swimmers.sc + #Demographic variables: sex, integument color, carapace width, and number of swimmers
    (1 | treatment.tank), #Use tank as a random effect to account for potential dependence between data
  data = modTTR,
  family = lognormal(link = "log") #Log normal model with a log link
)
```

```{r}
drop1(TTRmodelLogNorm) #Use a backwards deletion approach and AIC to identify best fit model. 

#Dropping the interaction term leads to the lowest AIC
```

```{r}
#Rerun the model without temperature:day interaction

TTRmodelLogNorm2 <- glmmTMB( #Generalized linear mixed effect model
  TTRavg ~ as.factor(treatment) + day.sc + #Treatment and day
    as.factor(sex) + integument.sc + carapace.width.sc + number.swimmers.sc + #Demographic variables: sex, integument color, carapace width, number of swimmers
    (1 | treatment.tank), #Use tank as a random effect to account for potential dependence between data
  data = modTTR,
  family = lognormal(link = "log") #Log normal model with a log link
)
```

```{r}
drop1(TTRmodelLogNorm2) #Use a backwards deletion approach and AIC to identify best fit model. 

#Dropping integument.sc will reduce AIC the most
```

```{r}
#Rerun the model without integument.sc

TTRmodelLogNorm3 <- glmmTMB( #Generalized linear mixed effect model
  TTRavg ~ as.factor(treatment) + day.sc + #Treatment and day
    as.factor(sex) + carapace.width.sc + number.swimmers.sc + #Demographic variables: sex, carapace width, number of swimmers
    (1 | treatment.tank), #Use tank as a random effect to account for potential dependence between data
  data = modTTR,
  family = lognormal(link = "log") #Log normal model with a log link
)
```

```{r}
drop1(TTRmodelLogNorm3) #Use a backwards deletion approach and AIC to identify best fit model. 

#Dropping day.sc will reduce AIC the most
```

```{r}
TTRmodelLogNorm4 <- glmmTMB( #Generalized linear mixed effect model
  TTRavg ~ as.factor(treatment) + #Treatment
    as.factor(sex) + carapace.width.sc + number.swimmers.sc + #Demographic variables: sex, carapace width, number of swimmers
    (1 | treatment.tank), #Use tank as a random effect to account for potential dependence between data
  data = modTTR,
  family = lognormal(link = "log") #Log normal model with a log link
)
```

```{r}
drop1(TTRmodelLogNorm4) #Use a backwards deletion approach and AIC to identify best fit model. 

#Removing sex will reduce AIC the most
```

```{r}
TTRmodelLogNorm5 <- glmmTMB( #Generalized linear mixed effect model
  TTRavg ~ as.factor(treatment) + #Treatment
    carapace.width.sc + number.swimmers.sc + #Demographic variables: carapace width, number of swimmers
    (1 | treatment.tank), #Use tank as a random effect to account for potential dependence between data
  data = modTTR,
  family = lognormal(link = "log") #Log normal model with a log link
)
```

```{r}
drop1(TTRmodelLogNorm5) #Use a backwards deletion approach and AIC to identify best fit model. 

#No more terms to remove!
```

### Inverse gaussian model family

```{r}
TTRmodelInvGaus <- glmmTMB( #Generalized linear mixed effect model
  TTRavg ~ as.factor(treatment) + day.sc + as.factor(treatment):day.sc + #Treatment, day, and their interaction
    as.factor(sex) + integument.sc + carapace.width.sc + number.swimmers.sc + #Demographic variables: sex, integument color, carapace width, and number of swimmers
    (1 | treatment.tank), #Use tank as a random effect to account for potential dependence between data
  data = modTTR,
  family = inverse.gaussian(link = "log") #Log normal model with a log link
)
```

```{r}
drop1(TTRmodelInvGaus) #Use a backwards deletion approach and AIC to identify best fit model. 

#Removing the interaction term leads to the lowest AIC
```

```{r}
TTRmodelInvGaus2 <- glmmTMB( #Generalized linear mixed effect model
  TTRavg ~ as.factor(treatment) + day.sc +  #Treatment, day
    as.factor(sex) + integument.sc + carapace.width.sc + number.swimmers.sc + #Demographic variables: sex, integument color, carapace width, and number of swimmers
    (1 | treatment.tank), #Use tank as a random effect to account for potential dependence between data
  data = modTTR,
  family = inverse.gaussian(link = "log") #Log normal model with a log link
)
```

```{r}
drop1(TTRmodelInvGaus2) #Use a backwards deletion approach and AIC to identify best fit model. 

#Removing sex leads to the lowest AIC
```

```{r}
TTRmodelInvGaus3 <- glmmTMB( #Generalized linear mixed effect model
  TTRavg ~ as.factor(treatment) + day.sc +  #Treatment, day
    integument.sc + carapace.width.sc + number.swimmers.sc + #Demographic variables: integument color, carapace width, and number of swimmers
    (1 | treatment.tank), #Use tank as a random effect to account for potential dependence between data
  data = modTTR,
  family = inverse.gaussian(link = "log") #Log normal model with a log link
)
```

```{r}
drop1(TTRmodelInvGaus3) #Use a backwards deletion approach and AIC to identify best fit model. 

#Removing number.swimmers.sc leads to the lowest AIC
```

```{r}
TTRmodelInvGaus4 <- glmmTMB( #Generalized linear mixed effect model
  TTRavg ~ as.factor(treatment) + day.sc +  #Treatment, day
    integument.sc + carapace.width.sc + #Demographic variables: integument color, carapace width
    (1 | treatment.tank), #Use tank as a random effect to account for potential dependence between data
  data = modTTR,
  family = inverse.gaussian(link = "log") #Log normal model with a log link
)
```

```{r}
drop1(TTRmodelInvGaus4) #Use a backwards deletion approach and AIC to identify best fit model. 

#No more terms to delete!
```

## Final model

```{r}
AIC(TTRmodelGamma4, TTRmodelGamma_disp4, TTRmodelLogNorm5, TTRmodelInvGaus4) #Compare AIC amongst the most parsimonious models for different model families
```
Since the gamma model has the lowest AIC, I will use that model as the final model.

```{r}
summary(TTRmodelGamma4) #Summary output of the final model
```

```{r}
Anova(TTRmodelGamma4, type = "III") %>% #Run type III Wald tests on final model
  broom::tidy(.) %>% #Tidy output
  write.csv(., "TTRmodelGamma4-typeIII-Wald-output.csv", quote = FALSE, row.names = FALSE) #Save output as a CSV file
```

## Check assumptions

### Multicollinearity

I want to ensure that my predictor variables are not correlated with eachother.

```{r}
check_collinearity(TTRmodelGamma4) # Check multicollinearity assumption.

#Adjusted VIF < 2, assumption is met, there is low correlation
```
### Heteroskedasticity

```{r}
sim_residuals <- simulateResiduals(fittedModel = TTRmodelGamma4, n = 1000, plot = TRUE) #Simulate residuals for the final model

#Residual plot is characteristic of underdispersion according to the DHARMa manual.
```

Based on the plot, there is slight non-uniformity in residual distribution. However, the distribution is almost linear, so that may be okay. There are no significant dispersion differences or any outliers. There is some curvature in the quantiles. 

```{r}
testUniformity(sim_residuals) #Kolmogorov-Smirnov test
```

### Dispersion

```{r}
testDispersion(sim_residuals) #Test dispersion

#Dispersion = 1.7266, but not significant overdispersion
```

### Outliers

```{r}
testOutliers(sim_residuals) #Test for outliers

#No significant outliers
```

### Quantile distributions

```{r}
testQuantiles(sim_residuals) #Test quantile distributions
```

```{r}
testCategorical(sim_residuals, catPred = ~ as.factor(treatment)) #Test quantile distributions against a specific categorial predictor variable

#Higher variance at 5ºC
```

### Normality of random effects

```{r}
re_normality <- check_normality(TTRmodelGamma4, effects = "random") # Check random effect normality
re_normality #Random effects are normally distributed (p = 0.762)

plot(re_normality) #QQ plot of random effects
```

## Model bootstrapping

Since the K-S test came out with a significant p-value and there are fluctuations in the quantile curves, I want to run a bootstrapping test on my model coefficients to confirm that my output is stable.

```{r}
boot_params <- bootstrap_parameters(TTRmodelGamma4, iterations = 1000) #Calculate bootstrapped p-values and 95% CI to on model coefficients
print(boot_params) #See results of boostrapping
```

```{r}
summary(TTRmodelGamma4)[[6]][[1]] #Original model output
```

Based on my comparison between the original model output and the bootstrapped p-values and confidence intervals, my original model inferences are stable!

## Save model output

```{r}
summary(TTRmodelGamma4)[[6]][[1]] %>% #Original model output
  as.data.frame(.) %>% #Convert to a dataframe
  rownames_to_column(., var = "Parameter") %>% #Convert rownames to Parameter column
  left_join(x = .,
            y = boot_params %>%
              as.data.frame(.), #Join with bootstrap output converted to a dataframe
            by = "Parameter") %>% #Join by the Parameter column
  write.csv(., "TTRmodelGamma4-output.csv", quote = FALSE, row.names = FALSE) #Save table version of GLMM output with boostrapping results
```

## Posthoc tests for significant predictors

There was a significant impacts of treatment, day, their interaction, and sex! I will investigate these effects by calculating estimated marginal means with `emmeans`.

### Treatment x day interaction

Since the interaction between treatment and day was significant, I need to examine treatment differences at each day.

```{r}
TTRmodel_EMM_tempDay <- emmeans(TTRmodelGamma4, #Model name
                            ~ treatment | day.sc, #Significant predictor to investigate
                            at = list(day.sc = scale(c(4, 8, 11, 15, 18, 22),
                                                     center = mean(modTTR$day),
                                                     scale = sd(modTTR$day))), #Days where TTR was measured. Include scale to automatically back-transform scaled predictor variable, and specify the center and scale variables used in the initial scaling for back-transformation.
                            type = "response", #Convert back to the response scale since a log link was used
                            adjust = "bonferroni") #Use Bonferroni correction for multiple comparisons
TTRmodel_EMM_tempDay #Estimated marginal means
```


```{r}
pairs(TTRmodel_EMM_tempDay) %>% #Get pairwise comparisons for treatment x day
  broom::tidy(.) %>% #Tidy output
  filter(., contrast != "30C / 5C") %>% #Remove comparisons between 30C and 5C
  mutate(., day = c(rep(4, times = 2),
                    rep(8, times = 2),
                    rep(11, times = 2),
                    rep(15, times = 2),
                    rep(18, times = 2),
                    rep(22, times = 2))) %>% #Create a column with actual date, not scaled date
  dplyr::select(day, contrast:adj.p.value) %>% #Retain columns of interest
  write.csv(., "TTRmodelGamma4-emmeans-tempDay.csv", quote = FALSE, row.names = FALSE) #Save as a csv
```

### Sex

```{r}
TTRmodel_EMM_sex <- emmeans(TTRmodelGamma4, #Model name
                            ~ sex, #Significant predictor to investigate
                            type = "response") #Convert back to the response scale since a log link was used
TTRmodel_EMM_sex #Estimated marginal means
```

```{r}
pairs(TTRmodel_EMM_sex) #Get pairwise comparisons
pairs(TTRmodel_EMM_sex) %>% #Take pairwise comparison output
  broom::tidy(.) %>% #Tidy it using broom so it is cleaner to look at
  write.csv("TTRmodelGamma4-emmeans-sex.csv", quote = FALSE, row.names = FALSE) #Save output as a csv.
```

```{r}
modTTR %>% #Formatted data
  group_by(sex) %>% #Group by sex
  summarize(.,avgTime = mean(TTRavg)) #Calculate average righting time for female and male crabs
```

```{r}
modTTR %>% #Formatted data
  ggplot(., mapping = aes(x = sex, y = TTRavg, color = sex)) + #Create a plot with sex on the x-axis, TTRavg on the y-axis, and colored by sex
  geom_boxplot() + geom_jitter(alpha = 0.5) + #Boxplot with jittered points and lighter colored points to see overlaps
  scale_x_discrete(name = "Sex") + #Relabel x-axis
  scale_y_continuous(name = "Average Time-to-Right (s)",
                     breaks = seq(0, 50, 10),
                     limits = c(0, 50)) + #Create y-axis
  theme_classic(base_size = 15) + theme(legend.position = "none") #Modify theme, include a base font size, and remove legend
ggsave("figures/TTRmodelGamma4-sex.pdf", height = 8.5, width = 11) #Save plot
```

# Threshold analysis

Another way to analyze this data is to count how many crabs within each treatment righted within a certain threshold. I can identify this threshold by taking the upper bound of a boxplot, then using a binomial model similar to [Coyle et al. (2019)](https://doi.org/10.1242/jeb.203521).

## Define threshold

```{r}
boxplot.stats(modTTR$TTRavg)$stats[5] #Use Tukey's method to identify upper bound of outliers
```

```{r}
modTTR <- modTTR %>%
  mutate(., TTRthresh = case_when(TTRavg > boxplot.stats(modTTR$TTRavg)$stats[5] ~ 0,
                                  TTRavg <= boxplot.stats(modTTR$TTRavg)$stats[5] ~ 1)) #Create a new column where 1 = righted within the upper bound of the boxplot, 0 = failure to right within that threshold
head(modTTR) #Confirm changes
```

## Binomial model

```{r}
TTRthreshModel <- glm(I(TTRthresh == 1) ~ treatment + day + treatment*day + as.factor(sex) + as.factor(integument.color) + carapace.width + weight + as.factor(missing.legs), 
                      family = binomial(), 
                      data = modTTR) #Create a binomial model, where 1 = success of righting wihtin the defined threshold
summary(TTRthreshModel) #Only the contrast bewteen 5C and 13C control is significant
```

```{r}
step(TTRthreshModel, test = "LRT") #Use likelihood ratio tests and backwards deletion approaches to identify the best model
```

## Final model

```{r}
TTRthreshModel2 <- glm(I(TTRthresh == 1) ~ treatment + day,
                       family = binomial(),
                       data = modTTR) #Run model selected by step with only significant terms
summary(TTRthreshModel2) #control vs. 5C significantly different, significant impact of day
```

## Save model output

```{r}
write.csv(broom::tidy(TTRthreshModel2), "TTR-binomialModel.csv", quote = FALSE, row.names = FALSE) #Save binomal GLM output
```

# Final plot

```{r}
modTTR %>%
  dplyr::select(., c(day, treatment, TTRavgFull, TTRavgFullLow, TTRavgFullHigh)) %>%
  distinct(.) %>%
  ggplot(., mapping = aes(x = day, y = TTRavgFull, color = treatment, shape = treatment)) +
  geom_pointrange(aes(ymin = TTRavgFullLow,
                      ymax = TTRavgFullHigh), 
                  size = 0.5, alpha = 0.75) +
  geom_line(y = boxplot.stats(modTTR$TTRavg)$stats[5], lty = 2, color = "black") +
  scale_x_continuous(name = "Day",
                     breaks = unique(modTTR$day),
                     limits = c(4, 22)) +
  scale_y_continuous(name = "Average Time-to-Right (s)",
                     breaks = c(0, seq(1, 5, 2), seq(5, 20, 5)),
                     limits = c(0, 17)) +
  scale_color_manual(values = c(plotColors[2], plotColors[1], plotColors[3]), 
                     name = "Temperature (ºC)",
                     breaks = c("13C", "30C", "5C"),
                     labels = c("13", "30", "5")) +
  scale_shape_manual(values = c(19, 17, 15),
                     name = "Temperature (ºC)",
                     breaks = c("13C", "30C", "5C"),
                     labels = c("13", "30", "5")) +
  theme_classic(base_size = 15) #Plot average TTR and add lm fits without SE bars. Add vertical line for boxplot outlier threshold. Modify x-axis to only show experimental days where TTR was measured. Scale y axis to include 91 = indication that crabs did not right within 90 s. Assign colors to each treatment. Increase base font size.
#ggsave("figures/time-to-right-avg.pdf", height = 8.5, width = 11)
```

```{r}
modTTR %>%
  dplyr::select(., c(day, treatment, TTRavg)) %>%
  distinct(.) %>%
  mutate(., treatment = gsub(x = treatment, pattern = "5C", replacement = "05C")) %>%
  mutate(., day = gsub(x = day, pattern = 4, replacement = "04")) %>%
  mutate(., day = gsub(x = day, pattern = 8, replacement = "08")) %>%
  mutate(., day = gsub(x = day, pattern = 11, replacement = "11")) %>%
  mutate(., day = gsub(x = day, pattern = 108, replacement = "18")) %>%
  mutate(., day = gsub(x = day, pattern = 15, replacement = "15")) %>%
  mutate(., day = gsub(x = day, pattern = 22, replacement = "22")) %>%
  ggplot(mapping = aes(x = day, y = TTRavg, color = treatment, shape = treatment)) +
  geom_boxplot(outlier.shape = NA) + geom_jitter(position = position_dodge(width = 0.75)) +
  # geom_hline(aes(yintercept = boxplot.stats(modTTR$TTRavg)$stats[5]), lty = 2, color = "black") +
  ylab("Average Time-to-Right (s)") +
  scale_x_discrete(name = "Day",
                   breaks = c("04", "08", "11", "18", "15", "22"),
                   labels = c("4", "8", "11", "18", "15", "22")) +
  scale_color_manual(values = c(plotColors[3], plotColors[2], plotColors[1]), 
                     name = "Temperature (ºC)",
                     breaks = c("05C", "13C", "30C"),
                     labels = c("5", "13", "30")) +
  scale_shape_manual(values = c(15, 19, 17),
                     name = "Temperature (ºC)",
                     breaks = c("05C", "13C", "30C"),
                     labels = c("5", "13", "30")) +
  theme_classic(base_size = 15) #Select unique day, treatment, and TTRavg data. Replace numbers with characters to facilitate better ordering. Plot average TTR for each crab in a boxplot. Do not show outliers with the OG boxplot, but add them in with geom_jitter. Add horizontal line for boxplot outlier threshold. Assign colors to each treatment. Increase base font size.
```

```{r}
TTRmodel_treatment$contrasts %>% broom::tidy(.) #Extract the contrast information to look at pairwise test output. Clean output using broom.
```


```{r}
mainPlot <- modTTR %>%
  dplyr::select(., c(day, treatment, TTRavg)) %>%
  distinct(.) %>%
  mutate(., treatment = gsub(x = treatment, pattern = "5C", replacement = "05C")) %>%
  mutate(., day = gsub(x = day, pattern = 4, replacement = "04")) %>%
  mutate(., day = gsub(x = day, pattern = 8, replacement = "08")) %>%
  mutate(., day = gsub(x = day, pattern = 11, replacement = "11")) %>%
  mutate(., day = gsub(x = day, pattern = 108, replacement = "18")) %>%
  mutate(., day = gsub(x = day, pattern = 15, replacement = "15")) %>%
  mutate(., day = gsub(x = day, pattern = 22, replacement = "22")) %>%
  ggplot(mapping = aes(x = day, y = TTRavg, color = treatment, shape = treatment)) +
  geom_boxplot(outlier.shape = NA) + geom_jitter(position = position_dodge(width = 0.75)) +
  ylab("Average Time-to-Right (s)") +
  scale_x_discrete(name = "Day",
                   breaks = c("04", "08", "11", "18", "15", "22"),
                   labels = c("4", "8", "11", "18", "15", "22")) +
  scale_color_manual(values = c(plotColors[3], plotColors[2], plotColors[1]), 
                     name = "Temperature (ºC)",
                     breaks = c("05C", "13C", "30C"),
                     labels = c("5", "13", "30")) +
  scale_shape_manual(values = c(15, 19, 17),
                     name = "Temperature (ºC)",
                     breaks = c("05C", "13C", "30C"),
                     labels = c("5", "13", "30")) +
  annotate(geom = "text", x = 7.5, y = 32,
           label = expression(paste(bolditalic("Temperature:  "), chi^2, " = 124.69, p < 0.0001")),
           hjust = 1) +
  annotate(geom = "text", x = 7.5, y = 30.5,
           label = expression(paste(bolditalic("Day:  "), chi^2, " = 11.21, p = 0.0008")),
           hjust = 1) +
  annotate(geom = "text", x = 7.5, y = 29,
           label = expression(paste(bolditalic("Temperature x Day:  "), chi^2, " = 12.15, p = 0.002")),
           hjust = 1) +
  annotate(geom = "text", x = 7.5, y = 27.5,
           label = expression(paste(bolditalic("Sex:  "), chi^2, " = 6.61, p = 0.01")),
           hjust = 1) +
  theme_classic(base_size = 15) #Select unique day, treatment, and TTRavg data. Replace numbers with characters to facilitate better ordering. Plot average TTR for each crab in a boxplot. Do not show outliers with the OG boxplot, but add them in with geom_jitter. Assign colors to each treatment. Add annotations with Wald test output for significant predictors. Increase base font size.

insetPlot <-  modTTR %>%
  dplyr::select(., c(day, treatment, TTRavg)) %>%
  distinct(.) %>%
  filter(., treatment != "5C") %>%
  mutate(., day = gsub(x = day, pattern = 4, replacement = "04")) %>%
  mutate(., day = gsub(x = day, pattern = 8, replacement = "08")) %>%
  mutate(., day = gsub(x = day, pattern = 11, replacement = "11")) %>%
  mutate(., day = gsub(x = day, pattern = 108, replacement = "18")) %>%
  mutate(., day = gsub(x = day, pattern = 15, replacement = "15")) %>%
  mutate(., day = gsub(x = day, pattern = 22, replacement = "22")) %>%
  ggplot(mapping = aes(x = day, y = TTRavg, color = treatment, shape = treatment)) +
  geom_boxplot(outlier.shape = NA) + geom_jitter(position = position_dodge(width = 0.75)) +
  ylab("") +
  scale_x_discrete(name = "",
                   breaks = c("04", "08", "11", "18", "15", "22"),
                   labels = c("4", "8", "11", "18", "15", "22")) +
  scale_color_manual(values = c(plotColors[2], plotColors[1]), 
                     name = "Temperature (ºC)",
                     breaks = c("13C", "30C"),
                     labels = c("13", "30")) +
  scale_shape_manual(values = c(19, 17),
                     name = "Temperature (ºC)",
                     breaks = c("13C", "30C"),
                     labels = c("13", "30")) +
  scale_y_continuous(limits = c(0,3)) +
  theme_classic() + theme(legend.position = "none",
                          axis.line.x.bottom = element_line(linewidth = 0),
                          axis.line.y.left = element_line(linewidth = 0),
                          panel.border = element_rect(colour = "black", fill = NA, size = 1))

plotWithInset <- ggdraw() +
  draw_plot(mainPlot) +
  draw_plot(insetPlot, x = 0.3, y = 0.73, width = 0.52, height = 0.25)

plotWithInset

ggsave("figures/time-to-right-avg-boxplot-noThresh.pdf", height = 8.5, width = 11)
```
