---
title: "04-respirometry-analysis"
author: "Yaamini Venkataraman"
date: '2023-04-10'
output: html_document
---

I will analyze the respirometry data from the experiment. I want to compare *C. maenas* rate of oxygen consumption between treatments and over time. 

# Set up R Markdown document

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = normalizePath("../output/04-respirometry-analysis/")) #Set root directory
```

```{r}
getwd()
```

# Install packages

```{r packages, include=FALSE}
#install.packages("tidyverse")
#install.packages("RColorBrewer")
#install.packages("respirometry")
require(tidyverse)
require(RColorBrewer)
require(respirometry)
```

```{r}
sessionInfo()
```

# Import data

At each time point, I have data for six tanks (three treatments, two tanks/treatment). Each tank has data for four probes, but only probes 1-3 were used to collect data. I will import data for each day in one file using `readr`, then ensure I have the correct columns.

```{r}
test <- read_delim("../../data/respiration/2022-07-26-resp/2022-07-26_093345_Tank_3/2022-07-26_093345_Tank_3.txt", skip = 72, delim = "\t", col_names = FALSE) %>% 
  select(., c(1:7, 12, 17, 21:25, 30, 35, 39:43, 48, 53, 57:61, 66, 71)) #Import data. Skip the first 72 rows
colnames(test) <- c("date", "time", 
                    "dt_s_1", "perc_air_sat_1", "dphi_1", "signal_int_mV_1", "amb_light_mV_1", "temp_C_1", "pressure_mbar_1",
                    "dt_s_2", "perc_air_sat_2", "dphi_2", "signal_int_mV_2", "amb_light_mV_2", "temp_C_2", "pressure_mbar_2",
                    "dt_s_3", "perc_air_sat_3", "dphi_3", "signal_int_mV_3", "amb_light_mV_3", "temp_C_3", "pressure_mbar_3",
                    "dt_s_4", "perc_air_sat_4", "dphi_4", "signal_int_mV_4", "amb_light_mV_4", "temp_C_4", "pressure_mbar_4") #Rename columns
head(test) #Confirm import
```

### 2022-07-08

```{r}
files20220708 <- fs::dir_ls("../../data/respiration/2022-07-08-resp/", glob = "*tank?.txt", recurse = TRUE) #Create a list of respirometry files from specified date. Use recurse = TRUE to prompt search through child directories. Only select files that end in tank[NUMBER].txt. ? = single character wildcard
files20220708 #Confirm the list has the correct files
```

```{r}
dat20220708 <- read_delim(files20220708, id = "tank", delim = "\t", skip = 72, col_names = FALSE)  %>% 
  select(., c(1:8, 13, 18, 22:26, 31, 36, 40:44, 49, 54)) #Import data from all files in the designated list. Specify tab-delimited files and skip the first 72 lines. There are no column names. Select data from probes 1-3 (probe 4 was neveer used. )
colnames(dat20220708) <- c("path", "date", "time", 
                    "dt_s_1", "perc_air_sat_1", "dphi_1", "signal_int_mV_1", "amb_light_mV_1", "temp_C_1", "pressure_mbar_1",
                    "dt_s_2", "perc_air_sat_2", "dphi_2", "signal_int_mV_2", "amb_light_mV_2", "temp_C_2", "pressure_mbar_2",
                    "dt_s_3", "perc_air_sat_3", "dphi_3", "signal_int_mV_3", "amb_light_mV_3", "temp_C_3", "pressure_mbar_3") #Rename columns
head(dat20220708)
```

```{r}
dat20220708 <- dat20220708 %>%
  mutate(., tank = str_sub(path, 82, 86)) %>%
  select(-path) %>% 
  mutate(., treatment = case_when(tank == "tank1" ~ "13C",
                                  tank == "tank2" ~ "5C",
                                  tank == "tank3" ~ "30C",
                                  tank == "tank4" ~ "13C",
                                  tank == "tank5" ~ "5C",
                                  tank == "tank6" ~ "30C")) #Isolate tank ID from the file path. Remove file path column. Add treatment information
tail(dat20220708) #Confirm tank and treatment information
```

### 2022-07-12

```{r}
files20220712 <- fs::dir_ls("../../data/respiration/2022-07-12-resp//", glob = "*tank?.txt", recurse = TRUE) #Create a list of respirometry files from specified date. Use recurse = TRUE to prompt search through child directories. Only select files that end in tank[NUMBER].txt. ? = single character wildcard
files20220712 #Confirm the list has the correct files
```

```{r}
dat20220712 <- read_delim(files20220712, id = "tank", delim = "\t", skip = 72, col_names = FALSE)  %>% 
  select(., c(1:8, 13, 18, 22:26, 31, 36, 40:44, 49, 54)) #Import data from all files in the designated list. Specify tab-delimited files and skip the first 72 lines. There are no column names.  Select data from probes 1-3 (probe 4 was neveer used. )
colnames(dat20220712) <- c("path", "date", "time", 
                    "dt_s_1", "perc_air_sat_1", "dphi_1", "signal_int_mV_1", "amb_light_mV_1", "temp_C_1", "pressure_mbar_1",
                    "dt_s_2", "perc_air_sat_2", "dphi_2", "signal_int_mV_2", "amb_light_mV_2", "temp_C_2", "pressure_mbar_2",
                    "dt_s_3", "perc_air_sat_3", "dphi_3", "signal_int_mV_3", "amb_light_mV_3", "temp_C_3", "pressure_mbar_3") #Rename columns
head(dat20220712)
```

```{r}
dat20220712 <- dat20220712 %>%
  mutate(., tank = str_sub(path, 82, 86)) %>%
  select(-path) %>% 
  mutate(., treatment = case_when(tank == "tank1" ~ "13C",
                                  tank == "tank2" ~ "5C",
                                  tank == "tank3" ~ "30C",
                                  tank == "tank4" ~ "13C",
                                  tank == "tank5" ~ "5C",
                                  tank == "tank6" ~ "30C")) #Isolate tank ID from the file path. Remove file path column. Add treatment information
tail(dat20220712) #Confirm tank and treatment information
```

### 2022-07-19

```{r}
files20220719 <- fs::dir_ls("../../data/respiration/2022-07-12-resp//", glob = "*tank?.txt", recurse = TRUE) #Create a list of respirometry files from specified date. Use recurse = TRUE to prompt search through child directories. Only select files that end in tank[NUMBER].txt. ? = single character wildcard
files20220719 #Confirm the list has the correct files
```

```{r}
dat20220719 <- read_delim(files20220719, id = "tank", delim = "\t", skip = 72, col_names = FALSE)  %>% 
  select(., c(1:8, 13, 18, 22:26, 31, 36, 40:44, 49, 54)) #Import data from all files in the designated list. Specify tab-delimited files and skip the first 72 lines. There are no column names. Select data from probes 1-3 (probe 4 was neveer used.)
colnames(dat20220719) <- c("path", "date", "time", 
                    "dt_s_1", "perc_air_sat_1", "dphi_1", "signal_int_mV_1", "amb_light_mV_1", "temp_C_1", "pressure_mbar_1",
                    "dt_s_2", "perc_air_sat_2", "dphi_2", "signal_int_mV_2", "amb_light_mV_2", "temp_C_2", "pressure_mbar_2",
                    "dt_s_3", "perc_air_sat_3", "dphi_3", "signal_int_mV_3", "amb_light_mV_3", "temp_C_3", "pressure_mbar_3") #Rename columns
head(dat20220719)
```

```{r}
dat20220719 <- dat20220719 %>%
  mutate(., tank = str_sub(path, 82, 86)) %>%
  select(-path) %>% 
  mutate(., treatment = case_when(tank == "tank1" ~ "13C",
                                  tank == "tank2" ~ "5C",
                                  tank == "tank3" ~ "30C",
                                  tank == "tank4" ~ "13C",
                                  tank == "tank5" ~ "5C",
                                  tank == "tank6" ~ "30C")) #Isolate tank ID from the file path. Remove file path column. Add treatment information
tail(dat20220719) #Confirm tank and treatment information
```

### 2022-07-26

```{r}
files20220726 <- fs::dir_ls("../../data/respiration/2022-07-12-resp//", glob = "*tank?.txt", recurse = TRUE) #Create a list of respirometry files from specified date. Use recurse = TRUE to prompt search through child directories. Only select files that end in tank[NUMBER].txt. ? = single character wildcard
files20220726 #Confirm the list has the correct files
```

```{r}
dat20220726 <- read_delim(files20220726, id = "tank", delim = "\t", skip = 72, col_names = FALSE)  %>% 
  select(., c(1:8, 13, 18, 22:26, 31, 36, 40:44, 49, 54)) #Import data from all files in the designated list. Specify tab-delimited files and skip the first 72 lines. There are no column names.
colnames(dat20220726) <- c("path", "date", "time", 
                    "dt_s_1", "perc_air_sat_1", "dphi_1", "signal_int_mV_1", "amb_light_mV_1", "temp_C_1", "pressure_mbar_1",
                    "dt_s_2", "perc_air_sat_2", "dphi_2", "signal_int_mV_2", "amb_light_mV_2", "temp_C_2", "pressure_mbar_2",
                    "dt_s_3", "perc_air_sat_3", "dphi_3", "signal_int_mV_3", "amb_light_mV_3", "temp_C_3", "pressure_mbar_3") #Rename columns
head(dat20220726)
```

```{r}
dat20220726 <- dat20220726 %>%
  mutate(., tank = str_sub(path, 82, 86)) %>%
  select(-path) %>% 
  mutate(., treatment = case_when(tank == "tank1" ~ "13C",
                                  tank == "tank2" ~ "5C",
                                  tank == "tank3" ~ "30C",
                                  tank == "tank4" ~ "13C",
                                  tank == "tank5" ~ "5C",
                                  tank == "tank6" ~ "30C")) #Isolate tank ID from the file path. Remove file path column. Add treatment information
tail(dat20220726) #Confirm tank and treatment information
```

# Format data


