---
title: "07-integration-analysis"
author: "Yaamini Venkataraman"
date: '2024-10-03'
output: html_document
---

In this script, I will integrate the metabolomics and lipidomics analyses. I will do this by correlating WCNA modules and using DIABLO to identify relationships between individual compounds.

# Set up R Markdown document

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = normalizePath("../output/07-integration-analysis/")) #Set root directory
```

```{r}
getwd()
```

# Install packages

```{r packages, warning = FALSE, message = FALSE}
# if ("tidyverse" %in% rownames(installed.packages()) == 'FALSE') install.packages('tidyverse')
# if ("plotrix" %in% rownames(installed.packages()) == 'FALSE') install.packages('plotrix')
# if ("vegan" %in% rownames(installed.packages()) == 'FALSE') install.packages('vegan')
# if ("ggfortify" %in% rownames(installed.packages()) == 'FALSE') install.packages('ggfortify')
# if (!requireNamespace("BiocManager", quietly = TRUE))
#   install.packages("BiocManager")
# if ("mixOmics" %in% rownames(installed.packages()) == 'FALSE') BiocManager::install("mixOmics")
# if ("RVAideMemoire" %in% rownames(installed.packages()) == 'FALSE') install.packages('RVAideMemoire')
# if ("broom" %in% rownames(installed.packages()) == 'FALSE') install.packages('broom')
# if ("RColorBrewer" %in% rownames(installed.packages()) == 'FALSE') install.packages('RColorBrewer')
# if ("pheatmap" %in% rownames(installed.packages()) == 'FALSE') install.packages('pheatmap')
# if ("genefilter" %in% rownames(installed.packages()) == 'FALSE') BiocManager::install("genefilter")
# if ("WGCNA" %in% rownames(installed.packages()) == 'FALSE') BiocManager::install('WGCNA')
# if ("dendsort" %in% rownames(installed.packages()) == 'FALSE') BiocManager::install('dendsort')
# if ("ComplexHeatmap" %in% rownames(installed.packages()) == 'FALSE') BiocManager::install('ComplexHeatmap')
# if ("svglite" %in% rownames(installed.packages()) == 'FALSE') install.packages("svglite")
require(tidyverse)
require(plotrix)
require(vegan)
require(ggfortify)
require(mixOmics)
require(RVAideMemoire)
require(broom)
require(RColorBrewer)
require(pheatmap)
require(genefilter)
require(WGCNA)
require(dendsort)
require(ComplexHeatmap)
require(svglite)
```

```{r}
sessionInfo()
```

```{bash}
mkdir figures
```

# Import and format data

## Metabolomics data

```{r}
attach("../05-metabolomics-analysis/05-metabolomics-analysis.RData") #Attach metabolomics RData
metabolomicsMetadata <- metabolomicsMetadata #Save metadata
transMetabData <- transMetabData #Save transposed data used for PLSDA
rowwiseMetabData <- rowwiseMetabData #Rowwise data used for WCNA
MEsMet <- MEs #Module eigengenes for metabolomics data
mME_metabolomics <- mME_meta #Data used to create module expression matrix
plotColors <- plotColors #Color scheme for plots
detach("file:../05-metabolomics-analysis/05-metabolomics-analysis.RData") #Detach metabolomics RData
```

```{r}
head(metabolomicsMetadata)
head(transMetabData)
head(rowwiseMetabData)
head(MEsMet)
head(mME_metabolomics)
plotColors
```

## Lipidomics data

```{r}
attach("../06-lipidomics-analysis/06-lipidomics-analysis.RData") #Attach lipidomics RData
lipidomicsMetadata <- lipidomicsMetadata #Save metadata
knownTransLipidData <- knownTransLipidData #Save transposed data used for PLSDA
rowwiseLipidData <- rowwiseLipidData #Rowwise data used for WCNA
mME_lipidomics <- mME_meta #Data used to create module expression matrix
detach("file:../06-lipidomics-analysis/06-lipidomics-analysis.RData") #Detach lipidomics RData
```

```{r}
head(lipidomicsMetadata)
head(knownTransLipidData)
head(rowwiseLipidData)
head(mME_lipidomics)
```

## Physiology data

### TTR data

```{r}
rawTTR <- read.csv("../../data/time-to-right.csv", header = TRUE) #Import raw data
head(rawTTR) #Confirm import. Trial data is in seconds.
```

```{r}
modTTR <- rawTTR %>%
  mutate(., day = case_when(date == "7/8/22" ~ 4,
                            date == "7/12/22" ~ 8,
                            date == "7/15/22" ~ 11,
                            date == "7/19/22" ~ 15,
                            date == "7/22/22" ~ 18,
                            date == "7/26/22" ~ 22)) %>%
  mutate(., treatment = case_when(treatment.tank == "1" | treatment.tank == "4" ~ "13C",
                                  treatment.tank == "2" | treatment.tank == "5" ~ "5C",
                                  treatment.tank == "3" | treatment.tank == "6" ~ "30C")) %>%
  mutate(., missing.legs = case_when(number.legs < 10 ~ "Y",
                                     number.legs == 10 ~ "N")) %>%
  mutate(., integument.cont = recode(integument.color,
                                     "B" = 0,
                                     "BG" = 0.5,
                                     "G" = 1, 
                                     "YG" = 1.5, 
                                     "Y" = 2, 
                                     "YO" = 2.5 , 
                                     "O" = 3, 
                                     "RO" = 3.5)) %>%
  mutate(., trial.1 = na_if(trial.1, 91.00)) %>%
  mutate(., trial.2 = na_if(trial.2, 91.00)) %>%
  mutate(., trial.3 = na_if(trial.3, 91.00)) %>%
  rowwise(.) %>% 
  mutate(., TTRavg = mean(c(trial.1, trial.2, trial.3), na.rm = TRUE)) %>%
  filter(., is.na(TTRavg) == FALSE) %>%
  mutate(., TTRSE = std.error(c(trial.1, trial.2, trial.3), na.rm = TRUE)) %>%
  filter(., is.na(TTRSE) == FALSE) %>%
  group_by(., day, treatment) %>%
  mutate(., TTRavgFull = mean(TTRavg)) %>%
  mutate(., TTRSEFull = std.error(TTRavg)) %>%
  mutate(., TTRavgFullLow = TTRavgFull - TTRSEFull) %>%
  mutate(., TTRavgFullHigh = TTRavgFull + TTRSEFull) %>%
  ungroup(.) #Remove sampling ID and notes columns. Add new column with day information. Add new column with treatment information. Create a a new column that with an index for width/length. Create new column as a binary for whether or not crabs are missing legs. Change all 91 to NA, calculate average and SE using data from the three TTR trials using rowwise operations, and remove rows where TTRavg | TTRSE = NA.  Calculate average and SE for all samples in a treatment for each day. Add/subtract TTRSEFull to/from TTRavgFull to get bounds. Ungroup.
head(modTTR) #Confirm formatting
```

I want to reduce the dataset down to crabs that were sampled. There were some crabs saved in different tubes as indicated in the notes column, but I fixed the tube labels when I sent them off for metabolomics sequencing!

```{r}
modTTRSampled <- modTTR %>%
  filter(is.na(sampling.ID) == FALSE) %>% 
  dplyr::select(-c(date, trial.1:trial.3, carapace.length, notes, TTRavgFull:TTRavgFullHigh)) %>% 
  rename(crab.ID = sampling.ID) %>%
  rename(tank = treatment.tank) %>%
  rename(probe_num = probe.number) %>% 
  mutate(treatment = gsub(pattern = "C", replacement = "", x = treatment) %>% as.factor(.)) %>%
  mutate(day = as.factor(day)) %>% 
  mutate(., unite(., "treatment_day", treatment:day, sep = "_")) %>%
  mutate(., treatment_day = gsub(x = treatment_day, pattern = "_  4", replacement = "_4")) %>%
  mutate(., treatment_day = gsub(x = treatment_day, pattern = "_ 22", replacement = "_22")) %>%
  mutate(., treatment_day = gsub(x = treatment_day, pattern = " 13_", replacement = "13_")) %>%
  mutate(., treatment_day = gsub(x = treatment_day, pattern = " 30_", replacement = "30_")) %>%
  mutate(., treatment_day = gsub(x = treatment_day, pattern = "  5_", replacement = "5_")) %>%
  mutate(., crab.ID = as.factor(crab.ID)) #Retain only crabs that were sampled and remove any superfluous columns. Rename specific columns to match respirometry data. Change certain columns to characters in order to match with metabolomics/lipidomics data.
head(modTTRSampled) #Confirm formatting
```

### Respirometry data

```{r}
attach("../04-respirometry-analysis/04-respirometry-analysis.RData") #Attach respirometry RData
all_MO2_slope_results_demTTRdata <- all_MO2_slope_results_demTTRdata #Data used for respirometry GLM
detach("file:../04-respirometry-analysis/04-respirometry-analysis.RData") #Detach respirometry RData
```

I want to match crab id with all of the demographic information in the respirometry dataframe.

```{r}
all_MO2_slope_results_sampled <- all_MO2_slope_results_demTTRdata %>%
  mutate(treatment = gsub(pattern = "C", replacement = "", x = treatment) %>% as.factor(.)) %>%
  mutate(., day = case_when(day == "04" ~ "4",
                            day != "04" ~ day) %>%
           as.factor(.)) %>%
  mutate(., unite(., "treatment_day", treatment:day, sep = "_")) %>%
  mutate(., treatment_day = gsub(x = treatment_day, pattern = "_  4", replacement = "_4")) %>%
  mutate(., treatment_day = gsub(x = treatment_day, pattern = "_ 22", replacement = "_22")) %>%
  mutate(., treatment_day = gsub(x = treatment_day, pattern = " 13_", replacement = "13_")) %>%
  mutate(., treatment_day = gsub(x = treatment_day, pattern = " 30_", replacement = "30_")) %>%
  mutate(., treatment_day = gsub(x = treatment_day, pattern = "  5_", replacement = "5_")) %>%
  filter(., day == "4" | day == "22") %>%
  left_join(., modTTRSampled, by = c("treatment", "day", "treatment_day", "tank", "probe_num", "sex", "integument.cont", "weight", "carapace.width", "missing.legs")) %>%
  dplyr::select(-c(TTRavg:TTRSE, integument.color, number.legs)) %>%
  mutate(., crab.ID = case_when(is.na(crab.ID) == TRUE ~ "304",
                                is.na(crab.ID) == FALSE ~ crab.ID)) #Change treatment, day, and treatment_day columns to match metabolomics/lipidomics datasets. Retain only respirometry data on the days that sampling was done. Join with modTTRSampled by all columns in common to get crab.ID. Remove superfluous columns. Manually change NA crab.ID to 304, since 304 was not included in the TTR data due to not righting.
head(all_MO2_slope_results_sampled) #Confirm formatting
```

# Correlating metabolomics and lipidomics data

To do this, I'll take the module expression information from each analysis and correlate them together!

```{r}
correlationInput <- left_join(mME_metabolomics %>%
                                pivot_wider(., names_prefix = "metabolite_", names_from = "Module", values_from = "value"),
                              mME_lipidomics %>%
                                pivot_wider(., names_prefix = "lipid_", names_from = "Module", values_from = "value"),
                              by = c("crab.ID")) %>%
  dplyr::select(., -c(treatment.y, day.y, treatment_day.y)) %>%
  dplyr::rename("treatment" = treatment.x, "day" = day.x, "treatment_day" = treatment_day.x) %>%
  dplyr::select(., -c("metabolite_1", "metabolite_8", "lipid_13", "lipid_14", "lipid_15"))
#Left join modified versions of module information tables from metabolomics and lipidomics data. Remove duplicate metadata information and rename columns to match standard naming conventions. Remove columns that were not significantly correlated to a treatment condition.
head(correlationInput)
```

```{r}
metabLipidCor <- cor(correlationInput[,5:24], method = "pearson") #Calculate Pearson's correlation coefficients
head(metabLipidCor %>% as.data.frame())
```

```{r}
nSamples <- nrow(rowwiseMetabData)
metabLipidPvalue <- corPvalueStudent(metabLipidCor, nSamples) #Calculate p-values for metabolite-lipid module correlations
head(metabLipidPvalue %>% as.data.frame())
```

```{r}
metabLipidCorClean <- as.data.frame(metabLipidCor) %>%
  dplyr::select(., starts_with("lipid_")) %>%
  filter(., row_number() %in% 1:7) #Take correlation data and keep columns where metabolite modules and correlated with lipid modules
head(metabLipidCorClean)
```

```{r}
metabLipidPvalueClean <- as.data.frame(metabLipidPvalue) %>%
  dplyr::select(., starts_with("lipid_")) %>%
  filter(., row_number() %in% 1:7) #Take p-values and keep columns where metabolite modules and correlated with lipid modules
head(metabLipidPvalueClean)
```

```{r}
#Save dataframes
write.csv(metabLipidCorClean, "metabolite-lipid-module-correlations.csv", quote = FALSE, row.names = TRUE)
write.csv(metabLipidPvalueClean, "metabolite-lipid-module-pvalues.csv", quote = FALSE, row.names = TRUE)
```


```{r, echo=TRUE, warning=FALSE, message=FALSE}
heatmappval <- signif(metabLipidPvalueClean, 1) #Create list of pvalues for eigengene correlation with specific traits

row_dend <- dendsort(hclust(dist(metabLipidCorClean))) #Create dendograms for rows
col_dend <- dendsort(hclust(dist(t(metabLipidCorClean)))) #Create dendograms for columns
```

```{r}
pdf(file = "figures/metabolite-lipid-module-relationship-heatmap.pdf", height = 8.5, width = 8.5)
ht <- Heatmap(as.matrix(metabLipidCorClean), name = "Eigengene", column_title = "Metabolomics-Lipidomics Eigengene Correlation",
              col = blueWhiteRed(50), 
              row_names_side = "left", 
              row_dend_side = "left",
              width = unit(5, "in"), 
              height = unit(4.5, "in"), 
              column_dend_reorder = TRUE, 
              cluster_columns = col_dend,
              row_dend_reorder = FALSE,  
              #column_split = 3, 
              #row_split = 3, 
              #column_dend_height = unit(.5, "in"),
              cluster_rows = row_dend, 
              #row_gap = unit(2.5, "mm"), 
              border = TRUE,
              cell_fun = function(j, i, x, y, w, h, col) {
                if(heatmappval[i, j] < 0.05) {
                  grid.text(sprintf("%s", heatmappval[i, j]), x, y, gp = gpar(fontsize = 10, fontface = "bold"))
                }
                else {
                  grid.text(sprintf("%s", heatmappval[i, j]), x, y, gp = gpar(fontsize = 10, fontface = "plain"))
                }},
              column_names_rot = 45,
              column_names_gp =  gpar(fontsize = 12, border = FALSE),
              column_title_gp = gpar(fontsize = 20),
              row_names_gp = gpar(fontsize = 12, alpha = 0.75, border = FALSE)) #Create a heatmap object
draw(ht) #Draw heatmap
dev.off()
```

### Repeat without module 0

Module 0 is where the WCNA algorithm places compounds that are not co-expressed with any other compounds. I want to recreate this heatmap without module 0, since I'm interested in co-expressed metabolites that are associated wtih co-expressed lipids to understand pathway-level processes.

```{r}
metabLipidCorClean <- as.data.frame(metabLipidCor) %>%
  dplyr::select(., starts_with("lipid_")) %>%
  filter(., row_number() %in% 1:7) #Take correlation data and keep columns where metabolite modules and correlated with lipid modules
head(metabLipidCorClean)
```

```{r}
metabLipidCorNo0 <- metabLipidCorClean %>%
  filter(., rownames(metabLipidCorClean) != "metabolite_0") %>%
  dplyr::select(-lipid_0) #Take dataframe and remove metabolite_0 row and lipid_0 column
head(metabLipidCorNo0) #Confirm changes
```

```{r}
metabLipidPvalueNo0 <- metabLipidPvalueClean %>%
  filter(., rownames(metabLipidPvalueClean) != "metabolite_0") %>%
  dplyr::select(-lipid_0) #Take dataframe and remove metabolite_0 row and lipid_0 column
head(metabLipidPvalueNo0) #Confirm changes
```

```{r}
#Save dataframes
write.csv(metabLipidCorNo0, "metabolite-lipid-module-correlations-no0.csv", quote = FALSE, row.names = TRUE)
write.csv(metabLipidPvalueNo0, "metabolite-lipid-module-pvalues-no0.csv", quote = FALSE, row.names = TRUE)
```


```{r, echo=TRUE, warning=FALSE, message=FALSE}
heatmappval <- signif(metabLipidPvalueNo0, 1) #Create list of pvalues for eigengene correlation with specific traits

row_dend <- dendsort(hclust(dist(metabLipidCorNo0))) #Create dendograms for rows
col_dend <- dendsort(hclust(dist(t(metabLipidCorNo0)))) #Create dendograms for columns
```

```{r}
#pdf(file = "figures/metabolite-lipid-module-relationship-heatmap-no0.pdf", height = 8.5, width = 8.5)
ht <- Heatmap(as.matrix(metabLipidCorNo0), name = "Eigengene", column_title = "Metabolomics-Lipidomics Eigengene Correlation",
              col = blueWhiteRed(50), 
              row_names_side = "left", 
              row_dend_side = "left",
              width = unit(5, "in"), 
              height = unit(4.5, "in"), 
              column_dend_reorder = TRUE, 
              cluster_columns = col_dend,
              row_dend_reorder = FALSE,  
              #column_split = 3, 
              #row_split = 3, 
              #column_dend_height = unit(.5, "in"),
              cluster_rows = row_dend, 
              #row_gap = unit(2.5, "mm"), 
              border = TRUE,
              cell_fun = function(j, i, x, y, w, h, col) {
                if(heatmappval[i, j] < 0.05) {
                  grid.text(sprintf("%s", heatmappval[i, j]), x, y, gp = gpar(fontsize = 10, fontface = "bold"))
                }
                else {
                  grid.text(sprintf("%s", heatmappval[i, j]), x, y, gp = gpar(fontsize = 10, fontface = "plain"))
                }},
              column_names_rot = 45,
              column_names_gp =  gpar(fontsize = 12, border = FALSE),
              column_title_gp = gpar(fontsize = 20),
              row_names_gp = gpar(fontsize = 12, alpha = 0.75, border = FALSE)) #Create a heatmap object
draw(ht) #Draw heatmap
#dev.off()
```

Removing these modules doesn't change the overall pattern! I will stick to this version without the null modules.

### Repeat with temperature-specific correlations

Now that I have established correlative relationships between metabolite and lipid modules, I want to understand how those relationships change with changing temperature. To do this, I want to examine these correlations for these same modules, but just using data for 30ºC, 13ºC, or 5ºC.

#### 13ºC

```{r}
metabLipidCor13 <- correlationInput %>%
  mutate(., treatment = gsub(pattern = " 13", replacement = "13", x = treatment)) %>%
  filter(., treatment == "13") %>%
  dplyr::select(., -c(1:4)) %>%
  cor(., method = "pearson") %>%
  as.data.frame(.) %>%
  dplyr::select(., starts_with("lipid_")) %>%
  filter(., row_number() %in% 1:7) %>%
  filter(., rownames(.) != "metabolite_0") %>%
  dplyr::select(-lipid_0) %>%
  as.matrix(.) #Take correlation input and filter data for 13ºC. Remove metadata columns and calculate Pearson's correlation coefficients. Convert to dataframe. Keep correlations where lipids are correlated with metabolites. Remove "0" modules. Convert back to a matrix
head(metabLipidCor13 %>% as.data.frame()) #Confirm formatting
```

```{r}
nSamples13 <- nrow(metabolomicsMetadata %>% filter(., treatment == " 13")) #Count the number of samples at 13ºC
metabLipidPvalue13 <- corPvalueStudent(metabLipidCor13, nSamples13) #Calculate p-values for metabolite-lipid module correlations
head(metabLipidPvalue13 %>% as.data.frame()) #Confirm output
```

```{r}
#Save dataframes
write.csv(metabLipidCor13, "metabolite-lipid-module-correlations-13.csv", quote = FALSE, row.names = TRUE)
write.csv(metabLipidPvalue13, "metabolite-lipid-module-pvalues-13.csv", quote = FALSE, row.names = TRUE)
```


```{r, echo=TRUE, warning=FALSE, message=FALSE}
heatmappval <- signif(metabLipidPvalue13, 1) #Create list of pvalues for eigengene correlation with specific traits

row_dend <- dendsort(hclust(dist(metabLipidCor13))) #Create dendograms for rows
col_dend <- dendsort(hclust(dist(t(metabLipidCor13)))) #Create dendograms for columns
```

```{r}
pdf(file = "figures/metabolite-lipid-module-relationship-heatmap-13.pdf", height = 8.5, width = 8.5)
ht <- Heatmap(as.matrix(metabLipidCor13), name = "Eigengene", column_title = "Metabolomics-Lipidomics Eigengene Correlation at 13ºC",
              col = blueWhiteRed(50), 
              row_names_side = "left", 
              row_dend_side = "left",
              width = unit(5, "in"), 
              height = unit(4.5, "in"), 
              column_dend_reorder = TRUE, 
              cluster_columns = col_dend,
              row_dend_reorder = FALSE,  
              #column_split = 3, 
              #row_split = 3, 
              #column_dend_height = unit(.5, "in"),
              cluster_rows = row_dend, 
              #row_gap = unit(2.5, "mm"), 
              border = TRUE,
              cell_fun = function(j, i, x, y, w, h, col) {
                if(heatmappval[i, j] < 0.05) {
                  grid.text(sprintf("%s", heatmappval[i, j]), x, y, gp = gpar(fontsize = 10, fontface = "bold"))
                }
                else {
                  grid.text(sprintf("%s", heatmappval[i, j]), x, y, gp = gpar(fontsize = 10, fontface = "plain"))
                }},
              column_names_rot = 45,
              column_names_gp =  gpar(fontsize = 12, border = FALSE),
              column_title_gp = gpar(fontsize = 20),
              row_names_gp = gpar(fontsize = 12, alpha = 0.75, border = FALSE)) #Create a heatmap object
draw(ht) #Draw heatmap
dev.off()
```

#### 30ºC

```{r}
metabLipidCor30 <- correlationInput %>%
  mutate(., treatment = gsub(pattern = " 30", replacement = "30", x = treatment)) %>%
  filter(., treatment == "30") %>%
  dplyr::select(., -c(1:4)) %>%
  cor(., method = "pearson") %>%
  as.data.frame(.) %>%
  dplyr::select(., starts_with("lipid_")) %>%
  filter(., row_number() %in% 1:7) %>%
  filter(., rownames(.) != "metabolite_0") %>%
  dplyr::select(-lipid_0) %>%
  as.matrix(.) #Take correlation input and filter data for 30ºC. Remove metadata columns and calculate Pearson's correlation coefficients. Convert to dataframe. Keep correlations where lipids are correlated with metabolites. Remove "0" modules. Convert back to a matrix
head(metabLipidCor30 %>% as.data.frame()) #Confirm formatting
```

```{r}
nSamples30 <- nrow(metabolomicsMetadata %>% filter(., treatment == " 30")) #Count the number of samples at 30ºC
metabLipidPvalue30 <- corPvalueStudent(metabLipidCor30, nSamples30) #Calculate p-values for metabolite-lipid module correlations
head(metabLipidPvalue30 %>% as.data.frame()) #Confirm output
```

```{r}
#Save dataframes
write.csv(metabLipidCor30, "metabolite-lipid-module-correlations-30.csv", quote = FALSE, row.names = TRUE)
write.csv(metabLipidPvalue30, "metabolite-lipid-module-pvalues-30.csv", quote = FALSE, row.names = TRUE)
```

```{r, echo=TRUE, warning=FALSE, message=FALSE}
heatmappval <- signif(metabLipidPvalue30, 1) #Create list of pvalues for eigengene correlation with specific traits

row_dend <- dendsort(hclust(dist(metabLipidCor30))) #Create dendograms for rows
col_dend <- dendsort(hclust(dist(t(metabLipidCor30)))) #Create dendograms for columns
```

```{r}
pdf(file = "figures/metabolite-lipid-module-relationship-heatmap-30.pdf", height = 8.5, width = 8.5)
ht <- Heatmap(as.matrix(metabLipidCor30), name = "Eigengene", column_title = "Metabolomics-Lipidomics Eigengene Correlation at 30ºC",
              col = blueWhiteRed(50), 
              row_names_side = "left", 
              row_dend_side = "left",
              width = unit(5, "in"), 
              height = unit(4.5, "in"), 
              column_dend_reorder = TRUE, 
              cluster_columns = col_dend,
              row_dend_reorder = FALSE,  
              #column_split = 3, 
              #row_split = 3, 
              #column_dend_height = unit(.5, "in"),
              cluster_rows = row_dend, 
              #row_gap = unit(2.5, "mm"), 
              border = TRUE,
              cell_fun = function(j, i, x, y, w, h, col) {
                if(heatmappval[i, j] < 0.05) {
                  grid.text(sprintf("%s", heatmappval[i, j]), x, y, gp = gpar(fontsize = 10, fontface = "bold"))
                }
                else {
                  grid.text(sprintf("%s", heatmappval[i, j]), x, y, gp = gpar(fontsize = 10, fontface = "plain"))
                }},
              column_names_rot = 45,
              column_names_gp =  gpar(fontsize = 12, border = FALSE),
              column_title_gp = gpar(fontsize = 20),
              row_names_gp = gpar(fontsize = 12, alpha = 0.75, border = FALSE)) #Create a heatmap object
draw(ht) #Draw heatmap
dev.off()
```

#### 5ºC

```{r}
metabLipidCor5 <- correlationInput %>%
  mutate(., treatment = gsub(pattern = "  5", replacement = "5", x = treatment)) %>%
  filter(., treatment == "5") %>%
  dplyr::select(., -c(1:4)) %>%
  cor(., method = "pearson") %>%
  as.data.frame(.) %>%
  dplyr::select(., starts_with("lipid_")) %>%
  filter(., row_number() %in% 1:7) %>%
  filter(., rownames(.) != "metabolite_0") %>%
  dplyr::select(-lipid_0) %>%
  as.matrix(.) #Take correlation input and filter data for 5ºC. Remove metadata columns and calculate Pearson's correlation coefficients. Convert to dataframe. Keep correlations where lipids are correlated with metabolites. Remove "0" modules. Convert back to a matrix
head(metabLipidCor5 %>% as.data.frame()) #Confirm formatting
```

```{r}
nSamples5 <- nrow(metabolomicsMetadata %>% filter(., treatment == "  5")) #Count the number of samples at 5ºC
metabLipidPvalue5 <- corPvalueStudent(metabLipidCor5, nSamples5) #Calculate p-values for metabolite-lipid module correlations
head(metabLipidPvalue5 %>% as.data.frame()) #Confirm output
```

```{r}
#Save dataframes
write.csv(metabLipidCor5, "metabolite-lipid-module-correlations-5.csv", quote = FALSE, row.names = TRUE)
write.csv(metabLipidPvalue5, "metabolite-lipid-module-pvalues-5.csv", quote = FALSE, row.names = TRUE)
```


```{r, echo=TRUE, warning=FALSE, message=FALSE}
heatmappval <- signif(metabLipidPvalue5, 1) #Create list of pvalues for eigengene correlation with specific traits

row_dend <- dendsort(hclust(dist(metabLipidCor5))) #Create dendograms for rows
col_dend <- dendsort(hclust(dist(t(metabLipidCor5)))) #Create dendograms for columns
```

```{r}
pdf(file = "figures/metabolite-lipid-module-relationship-heatmap-5.pdf", height = 8.5, width = 8.5)
ht <- Heatmap(as.matrix(metabLipidCor5), name = "Eigengene", column_title = "Metabolomics-Lipidomics Eigengene Correlation at 5ºC",
              col = blueWhiteRed(50), 
              row_names_side = "left", 
              row_dend_side = "left",
              width = unit(5, "in"), 
              height = unit(4.5, "in"), 
              column_dend_reorder = TRUE, 
              cluster_columns = col_dend,
              row_dend_reorder = FALSE,  
              #column_split = 3, 
              #row_split = 3, 
              #column_dend_height = unit(.5, "in"),
              cluster_rows = row_dend, 
              #row_gap = unit(2.5, "mm"), 
              border = TRUE,
              cell_fun = function(j, i, x, y, w, h, col) {
                if(heatmappval[i, j] < 0.05) {
                  grid.text(sprintf("%s", heatmappval[i, j]), x, y, gp = gpar(fontsize = 10, fontface = "bold"))
                }
                else {
                  grid.text(sprintf("%s", heatmappval[i, j]), x, y, gp = gpar(fontsize = 10, fontface = "plain"))
                }},
              column_names_rot = 45,
              column_names_gp =  gpar(fontsize = 12, border = FALSE),
              column_title_gp = gpar(fontsize = 20),
              row_names_gp = gpar(fontsize = 12, alpha = 0.75, border = FALSE)) #Create a heatmap object
draw(ht) #Draw heatmap
dev.off()
```

#### Create summary plot

My goal for this summary plot is to have the correlation coefficients for the metabolite x lipid modules across temperature. I will then be able to see if the correlation is primarily influenced by a specific temperature, or if it falls apart at a specific temperature.

```{r}
metabLipidCorNo0 %>%
  rownames_to_column(., var = "metabolite") %>%
  pivot_longer(cols = 2:13, names_to = "lipid", values_to = "correlation_overall") %>%
  left_join(., 
            (metabLipidPvalueNo0 %>%
               rownames_to_column(., var = "metabolite") %>%
               pivot_longer(cols = 2:13, names_to = "lipid", values_to = "pvalue_overall")),
            by = c("metabolite", "lipid")) %>%
  filter(., pvalue_overall < 0.05) %>%
  left_join(.,
            (metabLipidCor13 %>%
               as.data.frame(.) %>%
               rownames_to_column(., var = "metabolite") %>%
               pivot_longer(cols = 2:13, names_to = "lipid", values_to = "correlation_13") %>%
               left_join(., 
                         (metabLipidPvalue13 %>%
                            as.data.frame(.) %>%
                            rownames_to_column(., var = "metabolite") %>%
                            pivot_longer(cols = 2:13, names_to = "lipid", values_to = "pvalue_13")),
                         by = c("metabolite", "lipid"))),
            by = c("metabolite", "lipid")) %>%
  left_join(.,
            (metabLipidCor30 %>%
               as.data.frame(.) %>%
               rownames_to_column(., var = "metabolite") %>%
               pivot_longer(cols = 2:13, names_to = "lipid", values_to = "correlation_30") %>%
               left_join(., 
                         (metabLipidPvalue30 %>%
                            as.data.frame(.) %>%
                            rownames_to_column(., var = "metabolite") %>%
                            pivot_longer(cols = 2:13, names_to = "lipid", values_to = "pvalue_30")),
                         by = c("metabolite", "lipid"))),
            by = c("metabolite", "lipid")) %>%
  left_join(.,
            (metabLipidCor5 %>%
               as.data.frame(.) %>%
               rownames_to_column(., var = "metabolite") %>%
               pivot_longer(cols = 2:13, names_to = "lipid", values_to = "correlation_05") %>%
               left_join(., 
                         (metabLipidPvalue5 %>%
                            as.data.frame(.) %>%
                            rownames_to_column(., var = "metabolite") %>%
                            pivot_longer(cols = 2:13, names_to = "lipid", values_to = "pvalue_05")),
                         by = c("metabolite", "lipid"))),
            by = c("metabolite", "lipid")) %>%
  pivot_longer(.,
               cols = 3:10,
               names_to = c(".value", "condition"),
               names_pattern = "(.*)_(.*)") %>%
  ggplot(., aes(x = condition, y = correlation, color = condition, shape = condition)) +
  geom_point(size = 3) +
  facet_grid(metabolite ~ lipid, scales = "free_y") +
  labs(y = "Pearson's Correlation Coefficient") +
  scale_color_manual(name = "Condition",
                     values = c(rev(plotColors), "black"), 
                     labels = c("5ºC", "13ºC", "30ºC", "Overall")) +
  scale_shape_manual(values = c(15, 19, 17, 3),
                     name = "Condition",
                     labels = c("5ºC", "13ºC", "30ºC", "Overall")) +
  theme_classic(base_size = 15) + theme(axis.title.x = element_blank(),
                                        axis.ticks.x = element_blank(),
                                        axis.text.x = element_blank(),
                                        strip.background = element_rect(colour = "white")) #Take original correlation data and use a series of left_joins to merge with correlation data for each specific temperature that has already been pivoted longer. Finally, combine all correlation data into a single column and all p-value information into a single column using a specific regex pattern. Use information in a ggplot.
ggsave("figures/metabolte-lipid-module-relationship-correlations.pdf", height = 8.5, width = 11)
```

# Correlating physiology and -omics data

## WCNA

The next step is to add physiology information to the metabolomics and lipidomics WCNA analyses. I want to correlate metabolite and lipid profiles with average TTR and respirometry data on the days the crabs were dissected.

```{bash}
mkdir WCNA
mkdir WCNA/metabolomics
mkdir WCNA/lipidomics
```

### Modify trait matrix

```{r}
datTraitsMet <- metabolomicsMetadata %>%
  mutate(., day = gsub(x = day, pattern = 3, replacement = 4) %>% as.factor(.)) %>%
  mutate(., treatment_day = gsub(x = treatment_day, pattern = "_3", replacement = "_4") %>% as.factor(.)) %>%
  left_join(x = ., 
            y = (modTTRSampled %>% 
                   dplyr::select(c(crab.ID, treatment_day, TTRavg, TTRSE))), 
            by = c("crab.ID", "treatment_day")) %>%
  left_join(x = ., 
            y = all_MO2_slope_results_sampled %>%
              dplyr::select(c(crab.ID, treatment_day, slope_nmol_hr_corrected)), 
            by = c("crab.ID", "treatment_day")) %>%
  arrange(., crab.ID) %>%
  mutate(., "13ºC at Day 4" = case_when(treatment_day == "13_4" ~ 1,
                                        treatment_day != "13_4" ~ 0)) %>%
  mutate(., "5ºC at Day 4" = case_when(treatment_day == "5_4" ~ 1,
                                       treatment_day != "5_4" ~ 0)) %>%
  mutate(., "30ºC at Day 4" = case_when(treatment_day == "30_4" ~ 1,
                                        treatment_day != "30_4" ~ 0)) %>%
  mutate(., "13ºC at Day 22" = case_when(treatment_day == "13_22" ~ 1,
                                         treatment_day != "13_22" ~ 0)) %>%
  mutate(., "5ºC at Day 22" = case_when(treatment_day == "5_22" ~ 1,
                                        treatment_day != "5_22" ~ 0)) %>%
  mutate(., "30ºC at Day 22" = case_when(treatment_day == "30_22" ~ 1,
                                         treatment_day != "30_22" ~ 0)) %>%
  dplyr::select(-c(treatment:treatment_day, TTRSE)) %>%
  column_to_rownames(., var = "crab.ID") %>% 
  rename(., "Average TTR" = TTRavg) %>%
  rename(., "Oxygen Consumption" = slope_nmol_hr_corrected)
  #Take metadata and fix day to day 4 instead of day 3. Join with TTR data. 304, 308, and 506 do not have TTR data because they failed to right. Join with respirometry data (only select crabs have accompanying respirometry data). Use a series of case_when commands to create individual columns for each treatment or day, with 1 indicating that sample is part of the group, and 0 indicating that sample is not part of the group. Remove previous treatment classification columns and TTRSE column. Convert crab.ID to rownames. Rename TTR and respirometry columns
head(datTraitsMet) #Confirm dataframe creation
nrow(datTraitsMet) == nrow(rowwiseMetabData) #Confirm that there are the same number of rows (samples) between the new trait matrix and the metabolite data
```

### Correlate trait matrix with WCNA modules

```{r}
#Define numbers of metabolites and samples and view 
nMetabolites <- ncol(rowwiseMetabData)
nSamples <- nrow(rowwiseMetabData)

nMetabolites
nSamples
```

Next correlate the trait matrix (treatment, day) with eigengenes (modules).  

```{r}
moduleTraitCorMet <- cor(MEsMet, datTraitsMet, use = "p") #Correlations of traits with eigengenes
moduleTraitPvalueMet <- WGCNA::corPvalueStudent(moduleTraitCorMet, nSamples) #Get p-values for correlations
head(moduleTraitCorMet %>% as.data.frame())
head(moduleTraitPvalueMet %>% as.data.frame())
```

```{r}
moduleTraitTreeMet <- hclust(dist(t(moduleTraitCorMet)), method = "average") #Create cluster dendogram for module-trait correlations
```

```{r}
#pdf(file = "WCNA/figures/module-trait-cluster-dendogram.pdf", height = 8.5, width = 11)
plot(moduleTraitTreeMet) #Plot cluster dendogram
#dev.off()
```

### Plot correlations

```{bash}
mkdir WCNA/metabolomics/figures
```


```{r, echo=TRUE, warning=FALSE, message=FALSE}
textMatrixMet <- paste(signif(moduleTraitCorMet, 2), "\n(",signif(moduleTraitPvalueMet, 1), ")", sep = "") #Create a matrix with the module-trait correlation and p-values
dim(textMatrixMet) <- dim(moduleTraitCorMet) #Ensure this matrix has the appropriate dimensions
head(textMatrixMet) #Confirm matrix creation
```

```{r}
pdf("WCNA/metabolomics/figures/module-trait-relationships.pdf", height = 8.5, width = 11)
labeledHeatmap(Matrix = moduleTraitCorMet, xLabels = names(datTraitsMet),  yLabels = names(MEsMet), ySymbols = names(MEsMet), cex.lab.y = 1.5, cex.lab.x = 1.5, colors = blueWhiteRed(50), textMatrix = textMatrixMet, setStdMargins = TRUE, cex.text = 1, zlim = c(-1,1), main = paste("Module-trait relationships"), cex.main = 2) #Use the textMatrix to create a heatmap with correlations between modules and the various treatments.
dev.off()
```

```{r, echo=TRUE, warning=FALSE, message=FALSE}
heatmappval <- signif(moduleTraitPvalue, 1) #Create list of pvalues for eigengene correlation with specific traits

row_dend <- dendsort(hclust(dist(moduleTraitCor))) #Create dendograms for rows
col_dend <- dendsort(hclust(dist(t(moduleTraitCor)))) #Create dendograms for columns

trait_order <- c("5ºC at Day 3", "13ºC at Day 3", "30ºC at Day 3", "5ºC at Day 22", "13ºC at Day 22", "30ºC at Day 22") #Order of traits for heatmap
```

```{r}
#pdf(file = "WCNA/figures/module-trait-relationship-heatmap.pdf", height = 8.5, width = 8.5)
ht <- Heatmap(moduleTraitCor, name = "Eigengene", column_title = "Module-Group Eigengene Correlation",
              col = blueWhiteRed(50), 
              row_names_side = "left", 
              row_dend_side = "left",
              width = unit(5, "in"), 
              height = unit(4.5, "in"), 
              column_dend_reorder = TRUE, 
              cluster_columns = col_dend,
              row_dend_reorder = FALSE,  
              #column_split = 3, 
              #row_split = 3, 
              #column_dend_height = unit(.5, "in"),
              cluster_rows = row_dend, 
              #column_order = trait_order, 
              #row_gap = unit(2.5, "mm"), 
              border = TRUE,
              cell_fun = function(j, i, x, y, w, h, col) {
                if(heatmappval[i, j] < 0.05) {
                  grid.text(sprintf("%s", heatmappval[i, j]), x, y, gp = gpar(fontsize = 10, fontface = "bold"))
                }
                else {
                  grid.text(sprintf("%s", heatmappval[i, j]), x, y, gp = gpar(fontsize = 10, fontface = "plain"))
                }},
              column_names_rot = 45,
              column_names_gp =  gpar(fontsize = 12, border = FALSE),
              column_title_gp = gpar(fontsize = 20),
              row_names_gp = gpar(fontsize = 12, alpha = 0.75, border = FALSE)) #Create a heatmap object
draw(ht) #Draw heatmap
#dev.off()
```

## ASCA

## Individual metabolite and lipid correlations with DIABLO

Do correlations with VIP from significantly correlated modules?



