---
title: "07-integration-analysis"
author: "Yaamini Venkataraman"
date: '2024-10-03'
output: html_document
---

In this script, I will integrate the metabolomics and lipidomics analyses. I will do this by correlating WCNA modules and using DIABLO to identify relationships between individual compounds.

# Set up R Markdown document

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = normalizePath("../output/07-integration-analysis/")) #Set root directory
```

```{r}
getwd()
```

# Install packages

```{r packages, warning = FALSE, message = FALSE}
# if ("tidyverse" %in% rownames(installed.packages()) == 'FALSE') install.packages('tidyverse')
# if ("vegan" %in% rownames(installed.packages()) == 'FALSE') install.packages('vegan')
# if ("ggfortify" %in% rownames(installed.packages()) == 'FALSE') install.packages('ggfortify')
# if (!requireNamespace("BiocManager", quietly = TRUE))
#   install.packages("BiocManager")
# if ("mixOmics" %in% rownames(installed.packages()) == 'FALSE') BiocManager::install("mixOmics")
# if ("RVAideMemoire" %in% rownames(installed.packages()) == 'FALSE') install.packages('RVAideMemoire')
# if ("broom" %in% rownames(installed.packages()) == 'FALSE') install.packages('broom')
# if ("RColorBrewer" %in% rownames(installed.packages()) == 'FALSE') install.packages('RColorBrewer')
# if ("pheatmap" %in% rownames(installed.packages()) == 'FALSE') install.packages('pheatmap')
# if ("genefilter" %in% rownames(installed.packages()) == 'FALSE') BiocManager::install("genefilter")
# if ("WGCNA" %in% rownames(installed.packages()) == 'FALSE') BiocManager::install('WGCNA')
# if ("dendsort" %in% rownames(installed.packages()) == 'FALSE') BiocManager::install('dendsort')
# if ("ComplexHeatmap" %in% rownames(installed.packages()) == 'FALSE') BiocManager::install('ComplexHeatmap')
# if ("svglite" %in% rownames(installed.packages()) == 'FALSE') install.packages("svglite")
require(tidyverse)
require(vegan)
require(ggfortify)
require(mixOmics)
require(RVAideMemoire)
require(broom)
require(RColorBrewer)
require(pheatmap)
require(genefilter)
require(WGCNA)
require(dendsort)
require(ComplexHeatmap)
require(svglite)
```

```{r}
sessionInfo()
```

```{bash}
mkdir figures
```

# Import data

## Metabolomics data

```{r}
attach("../05-metabolomics-analysis/05-metabolomics-analysis.RData") #Attach metabolomics RData
metabolomicsMetadata <- metabolomicsMetadata #Save metadata
transMetabData <- transMetabData #Save transposed data used for PLSDA
rowwiseMetabData <- rowwiseMetabData #Rowwise data used for WCNA
mME_metabolomics <- mME_meta #Data used to create module expression matrix
detach("file:../05-metabolomics-analysis/05-metabolomics-analysis.RData") #Detach metabolomics RData
```

```{r}
head(metabolomicsMetadata)
head(transMetabData)
head(rowwiseMetabData)
head(mME_metabolomics)
```

## Lipidomics data

```{r}
attach("../06-lipidomics-analysis/06-lipidomics-analysis.RData") #Attach lipidomics RData
lipidomicsMetadata <- lipidomicsMetadata #Save metadata
knownTransLipidData <- knownTransLipidData #Save transposed data used for PLSDA
rowwiseLipidData <- rowwiseLipidData #Rowwise data used for WCNA
mME_lipidomics <- mME_meta #Data used to create module expression matrix
detach("file:../06-lipidomics-analysis/06-lipidomics-analysis.RData") #Detach lipidomics RData
```

```{r}
head(lipidomicsMetadata)
head(knownTransLipidData)
head(rowwiseLipidData)
head(mME_lipidomics)
```

## Physiology data

# Correlating physiology and -omics data

# Correlating metabolomics and lipidomics data

To do this, I'll take the module expression information from each analysis and correlate them together!

```{r}
correlationInput <- left_join(mME_metabolomics %>%
            pivot_wider(., names_prefix = "metabolite_", names_from = "Module", values_from = "value"),
          mME_lipidomics %>%
            pivot_wider(., names_prefix = "lipid_", names_from = "Module", values_from = "value"),
          by = c("crab.ID")) %>%
  dplyr::select(., -c(treatment.y, day.y, treatment_day.y)) %>%
  dplyr::rename("treatment" = treatment.x, "day" = day.x, "treatment_day" = treatment_day.x) %>%
  dplyr::select(., -c("metabolite_1", "metabolite_8", "lipid_13", "lipid_14", "lipid_15"))
  #Left join modified versions of module information tables from metabolomics and lipidomics data. Remove duplicate metadata information and rename columns to match standard naming conventions. Remove columns that were not significantly correlated to a treatment condition.
head(correlationInput)
```

```{r}
metabLipidCor <- cor(correlationInput[,5:24], method = "pearson") #Calculate Pearson's correlation coefficients
head(metabLipidCor %>% as.data.frame())
```

```{r}
nSamples <- nrow(rowwiseMetabData)
metabLipidPvalue <- corPvalueStudent(metabLipidCor, nSamples) #Calculate p-values for metabolite-lipid module correlations
head(metabLipidPvalue %>% as.data.frame())
```

```{r}
metabLipidCorClean <- as.data.frame(metabLipidCor) %>%
  dplyr::select(., starts_with("lipid_")) %>%
  filter(., row_number() %in% 1:7) #Take correlation data and keep columns where metabolite modules and correlated with lipid modules
head(metabLipidCorClean)
```

```{r}
metabLipidPvalueClean <- as.data.frame(metabLipidPvalue) %>%
  dplyr::select(., starts_with("lipid_")) %>%
  filter(., row_number() %in% 1:7) #Take p-values and keep columns where metabolite modules and correlated with lipid modules
head(metabLipidPvalueClean)
```

```{r}
#Save dataframes
write.csv(metabLipidCorClean, "metabolite-lipid-module-correlations.csv", quote = FALSE, row.names = TRUE)
write.csv(metabLipidPvalueClean, "metabolite-lipid-module-pvalues.csv", quote = FALSE, row.names = TRUE)
```


```{r, echo=TRUE, warning=FALSE, message=FALSE}
heatmappval <- signif(metabLipidPvalueClean, 1) #Create list of pvalues for eigengene correlation with specific traits

row_dend <- dendsort(hclust(dist(metabLipidCorClean))) #Create dendograms for rows
col_dend <- dendsort(hclust(dist(t(metabLipidCorClean)))) #Create dendograms for columns
```

```{r}
pdf(file = "figures/metabolite-lipid-module-relationship-heatmap.pdf", height = 8.5, width = 8.5)
ht <- Heatmap(as.matrix(metabLipidCorClean), name = "Eigengene", column_title = "Metabolomics-Lipidomics Eigengene Correlation",
              col = blueWhiteRed(50), 
              row_names_side = "left", 
              row_dend_side = "left",
              width = unit(5, "in"), 
              height = unit(4.5, "in"), 
              column_dend_reorder = TRUE, 
              cluster_columns = col_dend,
              row_dend_reorder = FALSE,  
              #column_split = 3, 
              #row_split = 3, 
              #column_dend_height = unit(.5, "in"),
              cluster_rows = row_dend, 
              #row_gap = unit(2.5, "mm"), 
              border = TRUE,
              cell_fun = function(j, i, x, y, w, h, col) {
                if(heatmappval[i, j] < 0.05) {
                  grid.text(sprintf("%s", heatmappval[i, j]), x, y, gp = gpar(fontsize = 10, fontface = "bold"))
                }
                else {
                  grid.text(sprintf("%s", heatmappval[i, j]), x, y, gp = gpar(fontsize = 10, fontface = "plain"))
                }},
              column_names_rot = 45,
              column_names_gp =  gpar(fontsize = 12, border = FALSE),
              column_title_gp = gpar(fontsize = 20),
              row_names_gp = gpar(fontsize = 12, alpha = 0.75, border = FALSE)) #Create a heatmap object
draw(ht) #Draw heatmap
dev.off()
```

### Repeat without module 0

Module 0 is where the WCNA algorithm places compounds that are not co-expressed with any other compounds. I want to recreate this heatmap without module 0, since I'm interested in co-expressed metabolites that are associated wtih co-expressed lipids to understand pathway-level processes.

```{r}
metabLipidCorNo0 <- as.data.frame(metabLipidCor) %>%
  dplyr::select(., starts_with("lipid_")) %>%
  filter(., row_number() %in% 1:7) #Take correlation data and keep columns where metabolite modules and correlated with lipid modules
head(metabLipidCorClean)
```

```{r}
metabLipidCorNo0 <- metabLipidCorClean %>%
  filter(., rownames(metabLipidCorClean) != "metabolite_0") %>%
  dplyr::select(-lipid_0) #Take dataframe and remove metabolite_0 row and lipid_0 column
head(metabLipidCorNo0) #Confirm changes
```

```{r}
metabLipidPvalueNo0 <- metabLipidPvalueClean %>%
  filter(., rownames(metabLipidPvalueClean) != "metabolite_0") %>%
  dplyr::select(-lipid_0) #Take dataframe and remove metabolite_0 row and lipid_0 column
head(metabLipidPvalueNo0) #Confirm changes
```

```{r}
#Save dataframes
write.csv(metabLipidCorNo0, "metabolite-lipid-module-correlations-no0.csv", quote = FALSE, row.names = TRUE)
write.csv(metabLipidPvalueNo0, "metabolite-lipid-module-pvalues-no0.csv", quote = FALSE, row.names = TRUE)
```


```{r, echo=TRUE, warning=FALSE, message=FALSE}
heatmappval <- signif(metabLipidPvalueClean, 1) #Create list of pvalues for eigengene correlation with specific traits

row_dend <- dendsort(hclust(dist(metabLipidCorClean))) #Create dendograms for rows
col_dend <- dendsort(hclust(dist(t(metabLipidCorClean)))) #Create dendograms for columns
```

```{r}
pdf(file = "figures/metabolite-lipid-module-relationship-heatmap.pdf", height = 8.5, width = 8.5)
ht <- Heatmap(as.matrix(metabLipidCorClean), name = "Eigengene", column_title = "Metabolomics-Lipidomics Eigengene Correlation",
              col = blueWhiteRed(50), 
              row_names_side = "left", 
              row_dend_side = "left",
              width = unit(5, "in"), 
              height = unit(4.5, "in"), 
              column_dend_reorder = TRUE, 
              cluster_columns = col_dend,
              row_dend_reorder = FALSE,  
              #column_split = 3, 
              #row_split = 3, 
              #column_dend_height = unit(.5, "in"),
              cluster_rows = row_dend, 
              #row_gap = unit(2.5, "mm"), 
              border = TRUE,
              cell_fun = function(j, i, x, y, w, h, col) {
                if(heatmappval[i, j] < 0.05) {
                  grid.text(sprintf("%s", heatmappval[i, j]), x, y, gp = gpar(fontsize = 10, fontface = "bold"))
                }
                else {
                  grid.text(sprintf("%s", heatmappval[i, j]), x, y, gp = gpar(fontsize = 10, fontface = "plain"))
                }},
              column_names_rot = 45,
              column_names_gp =  gpar(fontsize = 12, border = FALSE),
              column_title_gp = gpar(fontsize = 20),
              row_names_gp = gpar(fontsize = 12, alpha = 0.75, border = FALSE)) #Create a heatmap object
draw(ht) #Draw heatmap
dev.off()
```


# Individual metabolite and lipid correlations with DIABLO




