---
title: "07-integration-analysis"
author: "Yaamini Venkataraman"
date: '2024-10-03'
output: html_document
---

In this script, I will integrate the metabolomics and lipidomics analyses. I will do this by correlating WCNA modules and using DIABLO to identify relationships between individual compounds.

# Set up R Markdown document

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = normalizePath("../output/07-integration-analysis/")) #Set root directory
```

```{r}
getwd()
```

# Install packages

```{r packages, warning = FALSE, message = FALSE}
# if ("tidyverse" %in% rownames(installed.packages()) == 'FALSE') install.packages('tidyverse')
# if ("vegan" %in% rownames(installed.packages()) == 'FALSE') install.packages('vegan')
# if ("ggfortify" %in% rownames(installed.packages()) == 'FALSE') install.packages('ggfortify')
# if (!requireNamespace("BiocManager", quietly = TRUE))
#   install.packages("BiocManager")
# if ("mixOmics" %in% rownames(installed.packages()) == 'FALSE') BiocManager::install("mixOmics")
# if ("RVAideMemoire" %in% rownames(installed.packages()) == 'FALSE') install.packages('RVAideMemoire')
# if ("broom" %in% rownames(installed.packages()) == 'FALSE') install.packages('broom')
# if ("RColorBrewer" %in% rownames(installed.packages()) == 'FALSE') install.packages('RColorBrewer')
# if ("pheatmap" %in% rownames(installed.packages()) == 'FALSE') install.packages('pheatmap')
# if ("genefilter" %in% rownames(installed.packages()) == 'FALSE') BiocManager::install("genefilter")
# if ("WGCNA" %in% rownames(installed.packages()) == 'FALSE') BiocManager::install('WGCNA')
# if ("dendsort" %in% rownames(installed.packages()) == 'FALSE') BiocManager::install('dendsort')
# if ("ComplexHeatmap" %in% rownames(installed.packages()) == 'FALSE') BiocManager::install('ComplexHeatmap')
# if ("svglite" %in% rownames(installed.packages()) == 'FALSE') install.packages("svglite")
require(tidyverse)
require(vegan)
require(ggfortify)
require(mixOmics)
require(RVAideMemoire)
require(broom)
require(RColorBrewer)
require(pheatmap)
require(genefilter)
require(WGCNA)
require(dendsort)
require(ComplexHeatmap)
require(svglite)
```

```{r}
sessionInfo()
```

```{bash}
mkdir figures
```

# Import data

## Metabolomics data

```{r}
attach("../05-metabolomics-analysis/05-metabolomics-analysis.RData") #Attach metabolomics RData
metabolomicsMetadata <- metabolomicsMetadata #Save metadata
transMetabData <- transMetabData #Save transposed data used for PLSDA
rowwiseMetabData <- rowwiseMetabData #Rowwise data used for WCNA
mME_metabolomics <- mME_meta #Data used to create module expression matrix
detach("file:../05-metabolomics-analysis/05-metabolomics-analysis.RData") #Detach metabolomics RData
```

```{r}
head(metabolomicsMetadata)
head(transMetabData)
head(rowwiseMetabData)
head(mME_metabolomics)
```

## Lipidomics data

```{r}
attach("../06-lipidomics-analysis/06-lipidomics-analysis.RData") #Attach lipidomics RData
lipidomicsMetadata <- lipidomicsMetadata #Save metadata
knownTransLipidData <- knownTransLipidData #Save transposed data used for PLSDA
rowwiseLipidData <- rowwiseLipidData #Rowwise data used for WCNA
mME_lipidomics <- mME_meta #Data used to create module expression matrix
detach("file:../06-lipidomics-analysis/06-lipidomics-analysis.RData") #Detach lipidomics RData
```

```{r}
head(lipidomicsMetadata)
head(knownTransLipidData)
head(rowwiseLipidData)
head(mME_lipidomics)
```

## Physiology data

# Correlating physiology and -omics data

# Correlating metabolomics and lipidomics data

To do this, I'll take the module expression information from each analysis and correlate them together!

```{r}
correlationInput <- left_join(mME_metabolomics %>%
            pivot_wider(., names_prefix = "metabolite_", names_from = "Module", values_from = "value"),
          mME_lipidomics %>%
            pivot_wider(., names_prefix = "lipid_", names_from = "Module", values_from = "value"),
          by = c("crab.ID")) %>%
  dplyr::select(., -c(treatment.y, day.y, treatment_day.y)) %>%
  dplyr::rename("treatment" = treatment.x, "day" = day.x, "treatment_day" = treatment_day.x) %>%
  dplyr::select(., -c("metabolite_1", "metabolite_8", "lipid_13", "lipid_14", "lipid_15"))
  #Left join modified versions of module information tables from metabolomics and lipidomics data. Remove duplicate metadata information and rename columns to match standard naming conventions. Remove columns that were not significantly correlated to a treatment condition.
head(correlationInput)
```

```{r}
metabLipidCor <- cor(correlationInput[,5:24], method = "pearson") #Calculate Pearson's correlation coefficients
head(metabLipidCor %>% as.data.frame())
```

```{r}
nSamples <- nrow(rowwiseMetabData)
metabLipidPvalue <- corPvalueStudent(metabLipidCor, nSamples) #Calculate p-values for metabolite-lipid module correlations
head(metabLipidPvalue %>% as.data.frame())
```

```{r}
metabLipidCorClean <- as.data.frame(metabLipidCor) %>%
  dplyr::select(., starts_with("lipid_")) %>%
  filter(., row_number() %in% 1:7) #Take correlation data and keep columns where metabolite modules and correlated with lipid modules
head(metabLipidCorClean)
```

```{r}
metabLipidPvalueClean <- as.data.frame(metabLipidPvalue) %>%
  dplyr::select(., starts_with("lipid_")) %>%
  filter(., row_number() %in% 1:7) #Take p-values and keep columns where metabolite modules and correlated with lipid modules
head(metabLipidPvalueClean)
```

```{r}
#Save dataframes
write.csv(metabLipidCorClean, "metabolite-lipid-module-correlations.csv", quote = FALSE, row.names = TRUE)
write.csv(metabLipidPvalueClean, "metabolite-lipid-module-pvalues.csv", quote = FALSE, row.names = TRUE)
```


```{r, echo=TRUE, warning=FALSE, message=FALSE}
heatmappval <- signif(metabLipidPvalueClean, 1) #Create list of pvalues for eigengene correlation with specific traits

row_dend <- dendsort(hclust(dist(metabLipidCorClean))) #Create dendograms for rows
col_dend <- dendsort(hclust(dist(t(metabLipidCorClean)))) #Create dendograms for columns
```

```{r}
pdf(file = "figures/metabolite-lipid-module-relationship-heatmap.pdf", height = 8.5, width = 8.5)
ht <- Heatmap(as.matrix(metabLipidCorClean), name = "Eigengene", column_title = "Metabolomics-Lipidomics Eigengene Correlation",
              col = blueWhiteRed(50), 
              row_names_side = "left", 
              row_dend_side = "left",
              width = unit(5, "in"), 
              height = unit(4.5, "in"), 
              column_dend_reorder = TRUE, 
              cluster_columns = col_dend,
              row_dend_reorder = FALSE,  
              #column_split = 3, 
              #row_split = 3, 
              #column_dend_height = unit(.5, "in"),
              cluster_rows = row_dend, 
              #row_gap = unit(2.5, "mm"), 
              border = TRUE,
              cell_fun = function(j, i, x, y, w, h, col) {
                if(heatmappval[i, j] < 0.05) {
                  grid.text(sprintf("%s", heatmappval[i, j]), x, y, gp = gpar(fontsize = 10, fontface = "bold"))
                }
                else {
                  grid.text(sprintf("%s", heatmappval[i, j]), x, y, gp = gpar(fontsize = 10, fontface = "plain"))
                }},
              column_names_rot = 45,
              column_names_gp =  gpar(fontsize = 12, border = FALSE),
              column_title_gp = gpar(fontsize = 20),
              row_names_gp = gpar(fontsize = 12, alpha = 0.75, border = FALSE)) #Create a heatmap object
draw(ht) #Draw heatmap
dev.off()
```

### Repeat without module 0

Module 0 is where the WCNA algorithm places compounds that are not co-expressed with any other compounds. I want to recreate this heatmap without module 0, since I'm interested in co-expressed metabolites that are associated wtih co-expressed lipids to understand pathway-level processes.

```{r}
metabLipidCorClean <- as.data.frame(metabLipidCor) %>%
  dplyr::select(., starts_with("lipid_")) %>%
  filter(., row_number() %in% 1:7) #Take correlation data and keep columns where metabolite modules and correlated with lipid modules
head(metabLipidCorClean)
```

```{r}
metabLipidCorNo0 <- metabLipidCorClean %>%
  filter(., rownames(metabLipidCorClean) != "metabolite_0") %>%
  dplyr::select(-lipid_0) #Take dataframe and remove metabolite_0 row and lipid_0 column
head(metabLipidCorNo0) #Confirm changes
```

```{r}
metabLipidPvalueNo0 <- metabLipidPvalueClean %>%
  filter(., rownames(metabLipidPvalueClean) != "metabolite_0") %>%
  dplyr::select(-lipid_0) #Take dataframe and remove metabolite_0 row and lipid_0 column
head(metabLipidPvalueNo0) #Confirm changes
```

```{r}
#Save dataframes
write.csv(metabLipidCorNo0, "metabolite-lipid-module-correlations-no0.csv", quote = FALSE, row.names = TRUE)
write.csv(metabLipidPvalueNo0, "metabolite-lipid-module-pvalues-no0.csv", quote = FALSE, row.names = TRUE)
```


```{r, echo=TRUE, warning=FALSE, message=FALSE}
heatmappval <- signif(metabLipidPvalueNo0, 1) #Create list of pvalues for eigengene correlation with specific traits

row_dend <- dendsort(hclust(dist(metabLipidCorNo0))) #Create dendograms for rows
col_dend <- dendsort(hclust(dist(t(metabLipidCorNo0)))) #Create dendograms for columns
```

```{r}
#pdf(file = "figures/metabolite-lipid-module-relationship-heatmap-no0.pdf", height = 8.5, width = 8.5)
ht <- Heatmap(as.matrix(metabLipidCorNo0), name = "Eigengene", column_title = "Metabolomics-Lipidomics Eigengene Correlation",
              col = blueWhiteRed(50), 
              row_names_side = "left", 
              row_dend_side = "left",
              width = unit(5, "in"), 
              height = unit(4.5, "in"), 
              column_dend_reorder = TRUE, 
              cluster_columns = col_dend,
              row_dend_reorder = FALSE,  
              #column_split = 3, 
              #row_split = 3, 
              #column_dend_height = unit(.5, "in"),
              cluster_rows = row_dend, 
              #row_gap = unit(2.5, "mm"), 
              border = TRUE,
              cell_fun = function(j, i, x, y, w, h, col) {
                if(heatmappval[i, j] < 0.05) {
                  grid.text(sprintf("%s", heatmappval[i, j]), x, y, gp = gpar(fontsize = 10, fontface = "bold"))
                }
                else {
                  grid.text(sprintf("%s", heatmappval[i, j]), x, y, gp = gpar(fontsize = 10, fontface = "plain"))
                }},
              column_names_rot = 45,
              column_names_gp =  gpar(fontsize = 12, border = FALSE),
              column_title_gp = gpar(fontsize = 20),
              row_names_gp = gpar(fontsize = 12, alpha = 0.75, border = FALSE)) #Create a heatmap object
draw(ht) #Draw heatmap
#dev.off()
```

Removing these modules doesn't change the overall pattern! I will stick to this version without the null modules.

### Repeat with temperature-specific correlations

Now that I have established correlative relationships between metabolite and lipid modules, I want to understand how those relationships change with changing temperature. To do this, I want to examine these correlations for these same modules, but just using data for 30ºC, 13ºC, or 5ºC.

#### 13ºC

```{r}
metabLipidCor13 <- correlationInput %>%
  mutate(., treatment = gsub(pattern = " 13", replacement = "13", x = treatment)) %>%
  filter(., treatment == "13") %>%
  dplyr::select(., -c(1:4)) %>%
  cor(., method = "pearson") %>%
  as.data.frame(.) %>%
  dplyr::select(., starts_with("lipid_")) %>%
  filter(., row_number() %in% 1:7) %>%
  filter(., rownames(.) != "metabolite_0") %>%
  dplyr::select(-lipid_0) %>%
  as.matrix(.) #Take correlation input and filter data for 13ºC. Remove metadata columns and calculate Pearson's correlation coefficients. Convert to dataframe. Keep correlations where lipids are correlated with metabolites. Remove "0" modules. Convert back to a matrix
head(metabLipidCor13 %>% as.data.frame()) #Confirm formatting
```

```{r}
nSamples13 <- nrow(metabolomicsMetadata %>% filter(., treatment == " 13")) #Count the number of samples at 13ºC
metabLipidPvalue13 <- corPvalueStudent(metabLipidCor13, nSamples13) #Calculate p-values for metabolite-lipid module correlations
head(metabLipidPvalue13 %>% as.data.frame()) #Confirm output
```

```{r}
#Save dataframes
write.csv(metabLipidCor13, "metabolite-lipid-module-correlations-13.csv", quote = FALSE, row.names = TRUE)
write.csv(metabLipidPvalue13, "metabolite-lipid-module-pvalues-13.csv", quote = FALSE, row.names = TRUE)
```


```{r, echo=TRUE, warning=FALSE, message=FALSE}
heatmappval <- signif(metabLipidPvalue13, 1) #Create list of pvalues for eigengene correlation with specific traits

row_dend <- dendsort(hclust(dist(metabLipidCor13))) #Create dendograms for rows
col_dend <- dendsort(hclust(dist(t(metabLipidCor13)))) #Create dendograms for columns
```

```{r}
pdf(file = "figures/metabolite-lipid-module-relationship-heatmap-13.pdf", height = 8.5, width = 8.5)
ht <- Heatmap(as.matrix(metabLipidCor13), name = "Eigengene", column_title = "Metabolomics-Lipidomics Eigengene Correlation at 13ºC",
              col = blueWhiteRed(50), 
              row_names_side = "left", 
              row_dend_side = "left",
              width = unit(5, "in"), 
              height = unit(4.5, "in"), 
              column_dend_reorder = TRUE, 
              cluster_columns = col_dend,
              row_dend_reorder = FALSE,  
              #column_split = 3, 
              #row_split = 3, 
              #column_dend_height = unit(.5, "in"),
              cluster_rows = row_dend, 
              #row_gap = unit(2.5, "mm"), 
              border = TRUE,
              cell_fun = function(j, i, x, y, w, h, col) {
                if(heatmappval[i, j] < 0.05) {
                  grid.text(sprintf("%s", heatmappval[i, j]), x, y, gp = gpar(fontsize = 10, fontface = "bold"))
                }
                else {
                  grid.text(sprintf("%s", heatmappval[i, j]), x, y, gp = gpar(fontsize = 10, fontface = "plain"))
                }},
              column_names_rot = 45,
              column_names_gp =  gpar(fontsize = 12, border = FALSE),
              column_title_gp = gpar(fontsize = 20),
              row_names_gp = gpar(fontsize = 12, alpha = 0.75, border = FALSE)) #Create a heatmap object
draw(ht) #Draw heatmap
dev.off()
```

#### 30ºC

```{r}
metabLipidCor30 <- correlationInput %>%
  mutate(., treatment = gsub(pattern = " 30", replacement = "30", x = treatment)) %>%
  filter(., treatment == "30") %>%
  dplyr::select(., -c(1:4)) %>%
  cor(., method = "pearson") %>%
  as.data.frame(.) %>%
  dplyr::select(., starts_with("lipid_")) %>%
  filter(., row_number() %in% 1:7) %>%
  filter(., rownames(.) != "metabolite_0") %>%
  dplyr::select(-lipid_0) %>%
  as.matrix(.) #Take correlation input and filter data for 30ºC. Remove metadata columns and calculate Pearson's correlation coefficients. Convert to dataframe. Keep correlations where lipids are correlated with metabolites. Remove "0" modules. Convert back to a matrix
head(metabLipidCor30 %>% as.data.frame()) #Confirm formatting
```

```{r}
nSamples30 <- nrow(metabolomicsMetadata %>% filter(., treatment == " 30")) #Count the number of samples at 30ºC
metabLipidPvalue30 <- corPvalueStudent(metabLipidCor30, nSamples30) #Calculate p-values for metabolite-lipid module correlations
head(metabLipidPvalue30 %>% as.data.frame()) #Confirm output
```

```{r}
#Save dataframes
write.csv(metabLipidCor30, "metabolite-lipid-module-correlations-30.csv", quote = FALSE, row.names = TRUE)
write.csv(metabLipidPvalue30, "metabolite-lipid-module-pvalues-30.csv", quote = FALSE, row.names = TRUE)
```

```{r, echo=TRUE, warning=FALSE, message=FALSE}
heatmappval <- signif(metabLipidPvalue30, 1) #Create list of pvalues for eigengene correlation with specific traits

row_dend <- dendsort(hclust(dist(metabLipidCor30))) #Create dendograms for rows
col_dend <- dendsort(hclust(dist(t(metabLipidCor30)))) #Create dendograms for columns
```

```{r}
pdf(file = "figures/metabolite-lipid-module-relationship-heatmap-30.pdf", height = 8.5, width = 8.5)
ht <- Heatmap(as.matrix(metabLipidCor30), name = "Eigengene", column_title = "Metabolomics-Lipidomics Eigengene Correlation at 30ºC",
              col = blueWhiteRed(50), 
              row_names_side = "left", 
              row_dend_side = "left",
              width = unit(5, "in"), 
              height = unit(4.5, "in"), 
              column_dend_reorder = TRUE, 
              cluster_columns = col_dend,
              row_dend_reorder = FALSE,  
              #column_split = 3, 
              #row_split = 3, 
              #column_dend_height = unit(.5, "in"),
              cluster_rows = row_dend, 
              #row_gap = unit(2.5, "mm"), 
              border = TRUE,
              cell_fun = function(j, i, x, y, w, h, col) {
                if(heatmappval[i, j] < 0.05) {
                  grid.text(sprintf("%s", heatmappval[i, j]), x, y, gp = gpar(fontsize = 10, fontface = "bold"))
                }
                else {
                  grid.text(sprintf("%s", heatmappval[i, j]), x, y, gp = gpar(fontsize = 10, fontface = "plain"))
                }},
              column_names_rot = 45,
              column_names_gp =  gpar(fontsize = 12, border = FALSE),
              column_title_gp = gpar(fontsize = 20),
              row_names_gp = gpar(fontsize = 12, alpha = 0.75, border = FALSE)) #Create a heatmap object
draw(ht) #Draw heatmap
dev.off()
```

#### 5ºC

```{r}
metabLipidCor5 <- correlationInput %>%
  mutate(., treatment = gsub(pattern = "  5", replacement = "5", x = treatment)) %>%
  filter(., treatment == "5") %>%
  dplyr::select(., -c(1:4)) %>%
  cor(., method = "pearson") %>%
  as.data.frame(.) %>%
  dplyr::select(., starts_with("lipid_")) %>%
  filter(., row_number() %in% 1:7) %>%
  filter(., rownames(.) != "metabolite_0") %>%
  dplyr::select(-lipid_0) %>%
  as.matrix(.) #Take correlation input and filter data for 5ºC. Remove metadata columns and calculate Pearson's correlation coefficients. Convert to dataframe. Keep correlations where lipids are correlated with metabolites. Remove "0" modules. Convert back to a matrix
head(metabLipidCor5 %>% as.data.frame()) #Confirm formatting
```

```{r}
nSamples5 <- nrow(metabolomicsMetadata %>% filter(., treatment == "  5")) #Count the number of samples at 5ºC
metabLipidPvalue5 <- corPvalueStudent(metabLipidCor5, nSamples5) #Calculate p-values for metabolite-lipid module correlations
head(metabLipidPvalue5 %>% as.data.frame()) #Confirm output
```

```{r}
#Save dataframes
write.csv(metabLipidCor5, "metabolite-lipid-module-correlations-5.csv", quote = FALSE, row.names = TRUE)
write.csv(metabLipidPvalue5, "metabolite-lipid-module-pvalues-5.csv", quote = FALSE, row.names = TRUE)
```


```{r, echo=TRUE, warning=FALSE, message=FALSE}
heatmappval <- signif(metabLipidPvalue5, 1) #Create list of pvalues for eigengene correlation with specific traits

row_dend <- dendsort(hclust(dist(metabLipidCor5))) #Create dendograms for rows
col_dend <- dendsort(hclust(dist(t(metabLipidCor5)))) #Create dendograms for columns
```

```{r}
pdf(file = "figures/metabolite-lipid-module-relationship-heatmap-5.pdf", height = 8.5, width = 8.5)
ht <- Heatmap(as.matrix(metabLipidCor5), name = "Eigengene", column_title = "Metabolomics-Lipidomics Eigengene Correlation at 5ºC",
              col = blueWhiteRed(50), 
              row_names_side = "left", 
              row_dend_side = "left",
              width = unit(5, "in"), 
              height = unit(4.5, "in"), 
              column_dend_reorder = TRUE, 
              cluster_columns = col_dend,
              row_dend_reorder = FALSE,  
              #column_split = 3, 
              #row_split = 3, 
              #column_dend_height = unit(.5, "in"),
              cluster_rows = row_dend, 
              #row_gap = unit(2.5, "mm"), 
              border = TRUE,
              cell_fun = function(j, i, x, y, w, h, col) {
                if(heatmappval[i, j] < 0.05) {
                  grid.text(sprintf("%s", heatmappval[i, j]), x, y, gp = gpar(fontsize = 10, fontface = "bold"))
                }
                else {
                  grid.text(sprintf("%s", heatmappval[i, j]), x, y, gp = gpar(fontsize = 10, fontface = "plain"))
                }},
              column_names_rot = 45,
              column_names_gp =  gpar(fontsize = 12, border = FALSE),
              column_title_gp = gpar(fontsize = 20),
              row_names_gp = gpar(fontsize = 12, alpha = 0.75, border = FALSE)) #Create a heatmap object
draw(ht) #Draw heatmap
dev.off()
```

#### Create summary plot

### VIP in significantly correlated modules

do I use just these VIP in the DIABLO correlations??????

# Individual metabolite and lipid correlations with DIABLO


# TO-DO LIST:

- module eigengene correlation plot
  - overall module correlation vs. correlations with the data at specific temperatures
  - module expression at high temperature
  - module expressionat control temp
  - module expression at low temp?
- pull out VIP present in each module and plot abundance --> are there any a priori relationships to expect?




